### 2012-12-08 paper/paper.tex
i hope this is actually fixed

<!DOCTYPE html>
<html>
  <head prefix="og: http://ogp.me/ns# fb: http://ogp.me/ns/fb# githubog: http://ogp.me/ns/fb/githubog#">
    <meta charset='utf-8'>
    <meta http-equiv="X-UA-Compatible" content="IE=edge">
        <title>PSYCH204_Project/paper/paper.tex at master · ptbarthelemy/PSYCH204_Project</title>
    <link rel="search" type="application/opensearchdescription+xml" href="/opensearch.xml" title="GitHub" />
    <link rel="fluid-icon" href="https://github.com/fluidicon.png" title="GitHub" />
    <link rel="apple-touch-icon-precomposed" sizes="57x57" href="apple-touch-icon-114.png" />
    <link rel="apple-touch-icon-precomposed" sizes="114x114" href="apple-touch-icon-114.png" />
    <link rel="apple-touch-icon-precomposed" sizes="72x72" href="apple-touch-icon-144.png" />
    <link rel="apple-touch-icon-precomposed" sizes="144x144" href="apple-touch-icon-144.png" />
    <meta name="msapplication-TileImage" content="/windows-tile.png">
    <meta name="msapplication-TileColor" content="#ffffff">

    <link rel="icon" type="image/x-icon" href="/favicon.ico" />

    <meta content="authenticity_token" name="csrf-param" />
<meta content="WJuSSAjRdnVShD8r6DrOrsEgL/AxT6yLvH+B4+7oqpY=" name="csrf-token" />

    <link href="https://a248.e.akamai.net/assets.github.com/assets/github-a2a679dc329adac0ae0f3c8a55962cd292b6bb2e.css" media="screen" rel="stylesheet" type="text/css" />
    <link href="https://a248.e.akamai.net/assets.github.com/assets/github2-ddbc9d73271e204f76d48b6abb07e8791f16a0d5.css" media="screen" rel="stylesheet" type="text/css" />

    <script src="https://a248.e.akamai.net/assets.github.com/assets/frameworks-eee761b9d5e06efb064aaaf528c44ef8e1601e71.js" type="text/javascript"></script>
    <script src="https://a248.e.akamai.net/assets.github.com/assets/github-9dff60969c141f6060b5c427c34662424aefe349.js" type="text/javascript"></script>

      <link rel='permalink' href='/ptbarthelemy/PSYCH204_Project/blob/4f0d202c1ba791661a94da0e3af58abe72221df3/paper/paper.tex'>
    <meta property="og:title" content="PSYCH204_Project"/>
    <meta property="og:type" content="githubog:gitrepository"/>
    <meta property="og:url" content="https://github.com/ptbarthelemy/PSYCH204_Project"/>
    <meta property="og:image" content="https://secure.gravatar.com/avatar/2f3bac31ff97773344886e6ba59e1d49?s=420&amp;d=https://a248.e.akamai.net/assets.github.com%2Fimages%2Fgravatars%2Fgravatar-user-420.png"/>
    <meta property="og:site_name" content="GitHub"/>
    <meta property="og:description" content="Contribute to PSYCH204_Project development by creating an account on GitHub."/>

    <meta name="description" content="Contribute to PSYCH204_Project development by creating an account on GitHub." />

  <link href="https://github.com/ptbarthelemy/PSYCH204_Project/commits/master.atom" rel="alternate" title="Recent Commits to PSYCH204_Project:master" type="application/atom+xml" />

  </head>

  <body class="logged_in page-blob macintosh vis-public env-production ">
    <div id="wrapper">

        <div class="header header-logged-in true">
          <div class="container clearfix">

            <a class="header-logo-blacktocat" href="https://github.com/">
  <span class="mega-icon mega-icon-blacktocat"></span>
</a>

            <div class="divider-vertical"></div>

              <a href="/notifications" class="notification-indicator tooltipped downwards" title="You have no unread notifications">
    <span class="mail-status all-read"></span>
  </a>
  <div class="divider-vertical"></div>

  <div class="topsearch command-bar-activated">
    <form accept-charset="UTF-8" action="/search" class="command_bar_form" id="top_search_form" method="get">
  <a href="/search/advanced" class="advanced-search tooltipped downwards command-bar-search" id="advanced_search" title="Advanced search"><span class="mini-icon mini-icon-advanced-search "></span></a>

  <input type="text" name="q" id="command-bar" placeholder="Search or type a command" tabindex="1" data-username="nickborg" autocapitalize="off">

  <span class="mini-icon help tooltipped downwards" title="Show command bar help">
    <span class="mini-icon mini-icon-help"></span>
  </span>

  <input type="hidden" name="ref" value="commandbar">

  <div class="divider-vertical"></div>
</form>

    <ul class="top-nav">
        <li class="explore"><a href="https://github.com/explore">Explore</a></li>
        <li><a href="https://gist.github.com">Gist</a></li>
        <li><a href="/blog">Blog</a></li>
      <li><a href="http://help.github.com">Help</a></li>
    </ul>
  </div>

    <ul id="user-links">
      <li>
        <a href="https://github.com/nickborg" class="name">
          <img height="20" src="https://secure.gravatar.com/avatar/4c214ec0d06de350672ade1973246dd8?s=140&amp;d=https://a248.e.akamai.net/assets.github.com%2Fimages%2Fgravatars%2Fgravatar-user-420.png" width="20" /> nickborg
        </a>
      </li>
      <li>
        <a href="/new" id="new_repo" class="tooltipped downwards" title="Create a new repo">
          <span class="mini-icon mini-icon-create"></span>
        </a>
      </li>
      <li>
        <a href="/settings/profile" id="account_settings"
          class="tooltipped downwards"
          title="Account settings ">
          <span class="mini-icon mini-icon-account-settings"></span>
        </a>
      </li>
      <li>
          <a href="/logout" data-method="post" id="logout" class="tooltipped downwards" title="Sign out">
            <span class="mini-icon mini-icon-logout"></span>
          </a>
      </li>
    </ul>

          </div>
        </div>

            <div class="site hfeed" itemscope itemtype="http://schema.org/WebPage">
      <div class="hentry">

        <div class="pagehead repohead instapaper_ignore readability-menu">
          <div class="container">
            <div class="title-actions-bar">

                  <ul class="pagehead-actions">
          <li class="nspr">
            <a href="/ptbarthelemy/PSYCH204_Project/pull/new/master" class="minibutton btn-pull-request" icon_class="mini-icon-pull-request"><span class="mini-icon mini-icon-pull-request"></span>Pull Request</a>
          </li>

          <li class="subscription">
              <form accept-charset="UTF-8" action="/notifications/subscribe" data-autosubmit="true" data-remote="true" method="post"><div style="margin:0;padding:0;display:inline"><input name="authenticity_token" type="hidden" value="WJuSSAjRdnVShD8r6DrOrsEgL/AxT6yLvH+B4+7oqpY=" /></div>  <input id="repository_id" name="repository_id" type="hidden" value="6767293" />
  <div class="context-menu-container js-menu-container js-context-menu">
    <span class="minibutton switcher bigger js-menu-target">
      <span class="js-context-button">
          <span class="mini-icon mini-icon-unwatch"></span>Unwatch
      </span>
    </span>

    <div class="context-pane js-menu-content">
      <a href="#" class="close js-menu-close"><span class="mini-icon mini-icon-remove-close"></span></a>
      <div class="context-title">Notification status</div>

      <div class="context-body pane-selector">
        <ul class="js-navigation-container">
          <li class="selector-item js-navigation-item js-navigation-target ">
            <span class="mini-icon mini-icon-confirm"></span>
            <label>
              <input id="do_included" name="do" type="radio" value="included" />
              <h4>Not watching</h4>
              <p>You will only receive notifications when you participate or are mentioned.</p>
            </label>
            <span class="context-button-text js-context-button-text">
              <span class="mini-icon mini-icon-watching"></span>
              Watch
            </span>
          </li>
          <li class="selector-item js-navigation-item js-navigation-target selected">
            <span class="mini-icon mini-icon-confirm"></span>
            <label>
              <input checked="checked" id="do_subscribed" name="do" type="radio" value="subscribed" />
              <h4>Watching</h4>
              <p>You will receive all notifications for this repository.</p>
            </label>
            <span class="context-button-text js-context-button-text">
              <span class="mini-icon mini-icon-unwatch"></span>
              Unwatch
            </span>
          </li>
          <li class="selector-item js-navigation-item js-navigation-target ">
            <span class="mini-icon mini-icon-confirm"></span>
            <label>
              <input id="do_ignore" name="do" type="radio" value="ignore" />
              <h4>Ignored</h4>
              <p>You will not receive notifications for this repository.</p>
            </label>
            <span class="context-button-text js-context-button-text">
              <span class="mini-icon mini-icon-mute"></span>
              Stop ignoring
            </span>
          </li>
        </ul>
      </div>
    </div>
  </div>
</form>
          </li>

          <li class="js-toggler-container js-social-container starring-container ">
            <a href="/ptbarthelemy/PSYCH204_Project/unstar" class="minibutton js-toggler-target starred" data-remote="true" data-method="post" rel="nofollow">
              <span class="mini-icon mini-icon-star"></span>Unstar
            </a><a href="/ptbarthelemy/PSYCH204_Project/star" class="minibutton js-toggler-target unstarred" data-remote="true" data-method="post" rel="nofollow">
              <span class="mini-icon mini-icon-star"></span>Star
            </a><a class="social-count js-social-count" href="/ptbarthelemy/PSYCH204_Project/stargazers">0</a>
          </li>

              <li><a href="/ptbarthelemy/PSYCH204_Project/fork" class="minibutton js-toggler-target fork-button lighter" rel="nofollow" data-method="post"><span class="mini-icon mini-icon-fork"></span>Fork</a><a href="/ptbarthelemy/PSYCH204_Project/network" class="social-count">1</a>
              </li>

    </ul>

              <h1 itemscope itemtype="http://data-vocabulary.org/Breadcrumb" class="entry-title public">
                <span class="repo-label"><span>public</span></span>
                <span class="mega-icon mega-icon-public-repo"></span>
                <span class="author vcard">
                  <a href="/ptbarthelemy" class="url fn" itemprop="url" rel="author">
                  <span itemprop="title">ptbarthelemy</span>
                  </a></span> /
                <strong><a href="/ptbarthelemy/PSYCH204_Project" class="js-current-repository">PSYCH204_Project</a></strong>
              </h1>
            </div>

  <ul class="tabs">
    <li><a href="/ptbarthelemy/PSYCH204_Project" class="selected" highlight="repo_sourcerepo_downloadsrepo_commitsrepo_tagsrepo_branches">Code</a></li>
    <li><a href="/ptbarthelemy/PSYCH204_Project/network" highlight="repo_network">Network</a></li>
    <li><a href="/ptbarthelemy/PSYCH204_Project/pulls" highlight="repo_pulls">Pull Requests <span class='counter'>0</span></a></li>

      <li><a href="/ptbarthelemy/PSYCH204_Project/issues" highlight="repo_issues">Issues <span class='counter'>0</span></a></li>

      <li><a href="/ptbarthelemy/PSYCH204_Project/wiki" highlight="repo_wiki">Wiki</a></li>

    <li><a href="/ptbarthelemy/PSYCH204_Project/graphs" highlight="repo_graphsrepo_contributors">Graphs</a></li>

  </ul>

<div class="tabnav">

  <span class="tabnav-right">
    <ul class="tabnav-tabs">
      <li><a href="/ptbarthelemy/PSYCH204_Project/tags" class="tabnav-tab" highlight="repo_tags">Tags <span class="counter blank">0</span></a></li>
      <li><a href="/ptbarthelemy/PSYCH204_Project/downloads" class="tabnav-tab" highlight="repo_downloads">Downloads <span class="counter blank">0</span></a></li>
    </ul>

  </span>

  <div class="tabnav-widget scope">

    <div class="context-menu-container js-menu-container js-context-menu">
      <a href="#"
         class="minibutton bigger switcher js-menu-target js-commitish-button btn-branch repo-tree"
         data-hotkey="w"
         data-ref="master">
         <span><em class="mini-icon mini-icon-branch"></em><i>branch:</i> master</span>
      </a>

      <div class="context-pane commitish-context js-menu-content">
        <a href="#" class="close js-menu-close"><span class="mini-icon mini-icon-remove-close"></span></a>
        <div class="context-title">Switch branches/tags</div>
        <div class="context-body pane-selector commitish-selector js-navigation-container">
          <div class="filterbar">
            <input type="text" id="context-commitish-filter-field" class="js-navigation-enable js-filterable-field" placeholder="Filter branches/tags">
            <ul class="tabs">
              <li><a href="#" data-filter="branches" class="selected">Branches</a></li>
                <li><a href="#" data-filter="tags">Tags</a></li>
            </ul>
          </div>

          <div class="js-filter-tab js-filter-branches">
            <div data-filterable-for="context-commitish-filter-field" data-filterable-type=substring>
                <div class="commitish-item branch-commitish selector-item js-navigation-item js-navigation-target selected">
                  <span class="mini-icon mini-icon-confirm"></span>
                  <h4>
                      <a href="/ptbarthelemy/PSYCH204_Project/blob/master/paper/paper.tex" class="js-navigation-open" data-name="master" rel="nofollow">master</a>
                  </h4>
                </div>
            </div>
            <div class="no-results">Nothing to show</div>
          </div>

            <div class="js-filter-tab js-filter-tags filter-tab-empty" style="display:none">
              <div data-filterable-for="context-commitish-filter-field" data-filterable-type=substring>
              </div>
              <div class="no-results">Nothing to show</div>
            </div>
        </div>
      </div><!-- /.commitish-context-context -->
    </div>
  </div> <!-- /.scope -->

  <ul class="tabnav-tabs">
    <li><a href="/ptbarthelemy/PSYCH204_Project" class="selected tabnav-tab" highlight="repo_source">Files</a></li>
    <li><a href="/ptbarthelemy/PSYCH204_Project/commits/master" class="tabnav-tab" highlight="repo_commits">Commits</a></li>
    <li><a href="/ptbarthelemy/PSYCH204_Project/branches" class="tabnav-tab" highlight="repo_branches" rel="nofollow">Branches <span class="counter ">1</span></a></li>
  </ul>

</div>

          </div>
        </div><!-- /.repohead -->

        <div id="js-repo-pjax-container" class="container context-loader-container" data-pjax-container>

<!-- blob contrib key: blob_contributors:v21:2fd0485eaa95b9e1edea2e497c61e6f1 -->
<!-- blob contrib frag key: views10/v8/blob_contributors:v21:2fd0485eaa95b9e1edea2e497c61e6f1 -->

<div id="slider">

    <div class="frame-meta">

      <p title="This is a placeholder element" class="js-history-link-replace hidden"></p>
      <div class="breadcrumb">
        <span class='bold'><span itemscope="" itemtype="http://data-vocabulary.org/Breadcrumb"><a href="/ptbarthelemy/PSYCH204_Project" class="js-slide-to" data-direction="back" itemscope="url"><span itemprop="title">PSYCH204_Project</span></a></span></span> / <span itemscope="" itemtype="http://data-vocabulary.org/Breadcrumb"><a href="/ptbarthelemy/PSYCH204_Project/tree/master/paper" class="js-slide-to" data-direction="back" itemscope="url"><span itemprop="title">paper</span></a></span> / <strong class="final-path">paper.tex</strong> <span class="js-clippy mini-icon mini-icon-clippy " data-clipboard-text="paper/paper.tex" data-copied-hint="copied!" data-copy-hint="copy to clipboard"></span>
      </div>

      <a href="/ptbarthelemy/PSYCH204_Project/find/master" class="js-slide-to" data-hotkey="t" style="display:none">Show File Finder</a>

        <div class="commit commit-loader file-history-tease js-deferred-content" data-url="/ptbarthelemy/PSYCH204_Project/contributors/master/paper/paper.tex">
          Fetching contributors…

          <div class="participation">
            <p class="loader-loading"><img alt="Octocat-spinner-32-eaf2f5" height="16" src="https://a248.e.akamai.net/assets.github.com/images/spinners/octocat-spinner-32-EAF2F5.gif?1338945075" width="16" /></p>
            <p class="loader-error">Cannot retrieve contributors at this time</p>
          </div>
        </div>

    </div><!-- ./.frame-meta -->

    <div class="frames">
      <div class="frame" data-permalink-url="/ptbarthelemy/PSYCH204_Project/blob/4f0d202c1ba791661a94da0e3af58abe72221df3/paper/paper.tex" data-title="PSYCH204_Project/paper/paper.tex at master · ptbarthelemy/PSYCH204_Project · GitHub" data-type="blob">

        <div id="files" class="bubble">
          <div class="file">
            <div class="meta">
              <div class="info">
                <span class="icon"><b class="mini-icon mini-icon-text-file"></b></span>
                <span class="mode" title="File Mode">file</span>
                  <span>465 lines (347 sloc)</span>
                <span>30.783 kb</span>
              </div>
              <ul class="button-group actions">
                  <li>
                        <a class="grouped-button minibutton bigger lighter"
                           href="/ptbarthelemy/PSYCH204_Project/edit/master/paper/paper.tex"
                           data-method="post" rel="nofollow" data-hotkey="e">Edit</a>
                  </li>
                <li><a href="/ptbarthelemy/PSYCH204_Project/raw/master/paper/paper.tex" class="minibutton grouped-button bigger lighter" id="raw-url">Raw</a></li>
                  <li><a href="/ptbarthelemy/PSYCH204_Project/blame/master/paper/paper.tex" class="minibutton grouped-button bigger lighter">Blame</a></li>
                <li><a href="/ptbarthelemy/PSYCH204_Project/commits/master/paper/paper.tex" class="minibutton grouped-button bigger lighter" rel="nofollow">History</a></li>
              </ul>
            </div>
                <div class="data type-tex">
      <table cellpadding="0" cellspacing="0" class="lines">
        <tr>
          <td>
            <pre class="line_numbers"><span id="L1" rel="#L1">1</span>
<span id="L2" rel="#L2">2</span>
<span id="L3" rel="#L3">3</span>
<span id="L4" rel="#L4">4</span>
<span id="L5" rel="#L5">5</span>
<span id="L6" rel="#L6">6</span>
<span id="L7" rel="#L7">7</span>
<span id="L8" rel="#L8">8</span>
<span id="L9" rel="#L9">9</span>
<span id="L10" rel="#L10">10</span>
<span id="L11" rel="#L11">11</span>
<span id="L12" rel="#L12">12</span>
<span id="L13" rel="#L13">13</span>
<span id="L14" rel="#L14">14</span>
<span id="L15" rel="#L15">15</span>
<span id="L16" rel="#L16">16</span>
<span id="L17" rel="#L17">17</span>
<span id="L18" rel="#L18">18</span>
<span id="L19" rel="#L19">19</span>
<span id="L20" rel="#L20">20</span>
<span id="L21" rel="#L21">21</span>
<span id="L22" rel="#L22">22</span>
<span id="L23" rel="#L23">23</span>
<span id="L24" rel="#L24">24</span>
<span id="L25" rel="#L25">25</span>
<span id="L26" rel="#L26">26</span>
<span id="L27" rel="#L27">27</span>
<span id="L28" rel="#L28">28</span>
<span id="L29" rel="#L29">29</span>
<span id="L30" rel="#L30">30</span>
<span id="L31" rel="#L31">31</span>
<span id="L32" rel="#L32">32</span>
<span id="L33" rel="#L33">33</span>
<span id="L34" rel="#L34">34</span>
<span id="L35" rel="#L35">35</span>
<span id="L36" rel="#L36">36</span>
<span id="L37" rel="#L37">37</span>
<span id="L38" rel="#L38">38</span>
<span id="L39" rel="#L39">39</span>
<span id="L40" rel="#L40">40</span>
<span id="L41" rel="#L41">41</span>
<span id="L42" rel="#L42">42</span>
<span id="L43" rel="#L43">43</span>
<span id="L44" rel="#L44">44</span>
<span id="L45" rel="#L45">45</span>
<span id="L46" rel="#L46">46</span>
<span id="L47" rel="#L47">47</span>
<span id="L48" rel="#L48">48</span>
<span id="L49" rel="#L49">49</span>
<span id="L50" rel="#L50">50</span>
<span id="L51" rel="#L51">51</span>
<span id="L52" rel="#L52">52</span>
<span id="L53" rel="#L53">53</span>
<span id="L54" rel="#L54">54</span>
<span id="L55" rel="#L55">55</span>
<span id="L56" rel="#L56">56</span>
<span id="L57" rel="#L57">57</span>
<span id="L58" rel="#L58">58</span>
<span id="L59" rel="#L59">59</span>
<span id="L60" rel="#L60">60</span>
<span id="L61" rel="#L61">61</span>
<span id="L62" rel="#L62">62</span>
<span id="L63" rel="#L63">63</span>
<span id="L64" rel="#L64">64</span>
<span id="L65" rel="#L65">65</span>
<span id="L66" rel="#L66">66</span>
<span id="L67" rel="#L67">67</span>
<span id="L68" rel="#L68">68</span>
<span id="L69" rel="#L69">69</span>
<span id="L70" rel="#L70">70</span>
<span id="L71" rel="#L71">71</span>
<span id="L72" rel="#L72">72</span>
<span id="L73" rel="#L73">73</span>
<span id="L74" rel="#L74">74</span>
<span id="L75" rel="#L75">75</span>
<span id="L76" rel="#L76">76</span>
<span id="L77" rel="#L77">77</span>
<span id="L78" rel="#L78">78</span>
<span id="L79" rel="#L79">79</span>
<span id="L80" rel="#L80">80</span>
<span id="L81" rel="#L81">81</span>
<span id="L82" rel="#L82">82</span>
<span id="L83" rel="#L83">83</span>
<span id="L84" rel="#L84">84</span>
<span id="L85" rel="#L85">85</span>
<span id="L86" rel="#L86">86</span>
<span id="L87" rel="#L87">87</span>
<span id="L88" rel="#L88">88</span>
<span id="L89" rel="#L89">89</span>
<span id="L90" rel="#L90">90</span>
<span id="L91" rel="#L91">91</span>
<span id="L92" rel="#L92">92</span>
<span id="L93" rel="#L93">93</span>
<span id="L94" rel="#L94">94</span>
<span id="L95" rel="#L95">95</span>
<span id="L96" rel="#L96">96</span>
<span id="L97" rel="#L97">97</span>
<span id="L98" rel="#L98">98</span>
<span id="L99" rel="#L99">99</span>
<span id="L100" rel="#L100">100</span>
<span id="L101" rel="#L101">101</span>
<span id="L102" rel="#L102">102</span>
<span id="L103" rel="#L103">103</span>
<span id="L104" rel="#L104">104</span>
<span id="L105" rel="#L105">105</span>
<span id="L106" rel="#L106">106</span>
<span id="L107" rel="#L107">107</span>
<span id="L108" rel="#L108">108</span>
<span id="L109" rel="#L109">109</span>
<span id="L110" rel="#L110">110</span>
<span id="L111" rel="#L111">111</span>
<span id="L112" rel="#L112">112</span>
<span id="L113" rel="#L113">113</span>
<span id="L114" rel="#L114">114</span>
<span id="L115" rel="#L115">115</span>
<span id="L116" rel="#L116">116</span>
<span id="L117" rel="#L117">117</span>
<span id="L118" rel="#L118">118</span>
<span id="L119" rel="#L119">119</span>
<span id="L120" rel="#L120">120</span>
<span id="L121" rel="#L121">121</span>
<span id="L122" rel="#L122">122</span>
<span id="L123" rel="#L123">123</span>
<span id="L124" rel="#L124">124</span>
<span id="L125" rel="#L125">125</span>
<span id="L126" rel="#L126">126</span>
<span id="L127" rel="#L127">127</span>
<span id="L128" rel="#L128">128</span>
<span id="L129" rel="#L129">129</span>
<span id="L130" rel="#L130">130</span>
<span id="L131" rel="#L131">131</span>
<span id="L132" rel="#L132">132</span>
<span id="L133" rel="#L133">133</span>
<span id="L134" rel="#L134">134</span>
<span id="L135" rel="#L135">135</span>
<span id="L136" rel="#L136">136</span>
<span id="L137" rel="#L137">137</span>
<span id="L138" rel="#L138">138</span>
<span id="L139" rel="#L139">139</span>
<span id="L140" rel="#L140">140</span>
<span id="L141" rel="#L141">141</span>
<span id="L142" rel="#L142">142</span>
<span id="L143" rel="#L143">143</span>
<span id="L144" rel="#L144">144</span>
<span id="L145" rel="#L145">145</span>
<span id="L146" rel="#L146">146</span>
<span id="L147" rel="#L147">147</span>
<span id="L148" rel="#L148">148</span>
<span id="L149" rel="#L149">149</span>
<span id="L150" rel="#L150">150</span>
<span id="L151" rel="#L151">151</span>
<span id="L152" rel="#L152">152</span>
<span id="L153" rel="#L153">153</span>
<span id="L154" rel="#L154">154</span>
<span id="L155" rel="#L155">155</span>
<span id="L156" rel="#L156">156</span>
<span id="L157" rel="#L157">157</span>
<span id="L158" rel="#L158">158</span>
<span id="L159" rel="#L159">159</span>
<span id="L160" rel="#L160">160</span>
<span id="L161" rel="#L161">161</span>
<span id="L162" rel="#L162">162</span>
<span id="L163" rel="#L163">163</span>
<span id="L164" rel="#L164">164</span>
<span id="L165" rel="#L165">165</span>
<span id="L166" rel="#L166">166</span>
<span id="L167" rel="#L167">167</span>
<span id="L168" rel="#L168">168</span>
<span id="L169" rel="#L169">169</span>
<span id="L170" rel="#L170">170</span>
<span id="L171" rel="#L171">171</span>
<span id="L172" rel="#L172">172</span>
<span id="L173" rel="#L173">173</span>
<span id="L174" rel="#L174">174</span>
<span id="L175" rel="#L175">175</span>
<span id="L176" rel="#L176">176</span>
<span id="L177" rel="#L177">177</span>
<span id="L178" rel="#L178">178</span>
<span id="L179" rel="#L179">179</span>
<span id="L180" rel="#L180">180</span>
<span id="L181" rel="#L181">181</span>
<span id="L182" rel="#L182">182</span>
<span id="L183" rel="#L183">183</span>
<span id="L184" rel="#L184">184</span>
<span id="L185" rel="#L185">185</span>
<span id="L186" rel="#L186">186</span>
<span id="L187" rel="#L187">187</span>
<span id="L188" rel="#L188">188</span>
<span id="L189" rel="#L189">189</span>
<span id="L190" rel="#L190">190</span>
<span id="L191" rel="#L191">191</span>
<span id="L192" rel="#L192">192</span>
<span id="L193" rel="#L193">193</span>
<span id="L194" rel="#L194">194</span>
<span id="L195" rel="#L195">195</span>
<span id="L196" rel="#L196">196</span>
<span id="L197" rel="#L197">197</span>
<span id="L198" rel="#L198">198</span>
<span id="L199" rel="#L199">199</span>
<span id="L200" rel="#L200">200</span>
<span id="L201" rel="#L201">201</span>
<span id="L202" rel="#L202">202</span>
<span id="L203" rel="#L203">203</span>
<span id="L204" rel="#L204">204</span>
<span id="L205" rel="#L205">205</span>
<span id="L206" rel="#L206">206</span>
<span id="L207" rel="#L207">207</span>
<span id="L208" rel="#L208">208</span>
<span id="L209" rel="#L209">209</span>
<span id="L210" rel="#L210">210</span>
<span id="L211" rel="#L211">211</span>
<span id="L212" rel="#L212">212</span>
<span id="L213" rel="#L213">213</span>
<span id="L214" rel="#L214">214</span>
<span id="L215" rel="#L215">215</span>
<span id="L216" rel="#L216">216</span>
<span id="L217" rel="#L217">217</span>
<span id="L218" rel="#L218">218</span>
<span id="L219" rel="#L219">219</span>
<span id="L220" rel="#L220">220</span>
<span id="L221" rel="#L221">221</span>
<span id="L222" rel="#L222">222</span>
<span id="L223" rel="#L223">223</span>
<span id="L224" rel="#L224">224</span>
<span id="L225" rel="#L225">225</span>
<span id="L226" rel="#L226">226</span>
<span id="L227" rel="#L227">227</span>
<span id="L228" rel="#L228">228</span>
<span id="L229" rel="#L229">229</span>
<span id="L230" rel="#L230">230</span>
<span id="L231" rel="#L231">231</span>
<span id="L232" rel="#L232">232</span>
<span id="L233" rel="#L233">233</span>
<span id="L234" rel="#L234">234</span>
<span id="L235" rel="#L235">235</span>
<span id="L236" rel="#L236">236</span>
<span id="L237" rel="#L237">237</span>
<span id="L238" rel="#L238">238</span>
<span id="L239" rel="#L239">239</span>
<span id="L240" rel="#L240">240</span>
<span id="L241" rel="#L241">241</span>
<span id="L242" rel="#L242">242</span>
<span id="L243" rel="#L243">243</span>
<span id="L244" rel="#L244">244</span>
<span id="L245" rel="#L245">245</span>
<span id="L246" rel="#L246">246</span>
<span id="L247" rel="#L247">247</span>
<span id="L248" rel="#L248">248</span>
<span id="L249" rel="#L249">249</span>
<span id="L250" rel="#L250">250</span>
<span id="L251" rel="#L251">251</span>
<span id="L252" rel="#L252">252</span>
<span id="L253" rel="#L253">253</span>
<span id="L254" rel="#L254">254</span>
<span id="L255" rel="#L255">255</span>
<span id="L256" rel="#L256">256</span>
<span id="L257" rel="#L257">257</span>
<span id="L258" rel="#L258">258</span>
<span id="L259" rel="#L259">259</span>
<span id="L260" rel="#L260">260</span>
<span id="L261" rel="#L261">261</span>
<span id="L262" rel="#L262">262</span>
<span id="L263" rel="#L263">263</span>
<span id="L264" rel="#L264">264</span>
<span id="L265" rel="#L265">265</span>
<span id="L266" rel="#L266">266</span>
<span id="L267" rel="#L267">267</span>
<span id="L268" rel="#L268">268</span>
<span id="L269" rel="#L269">269</span>
<span id="L270" rel="#L270">270</span>
<span id="L271" rel="#L271">271</span>
<span id="L272" rel="#L272">272</span>
<span id="L273" rel="#L273">273</span>
<span id="L274" rel="#L274">274</span>
<span id="L275" rel="#L275">275</span>
<span id="L276" rel="#L276">276</span>
<span id="L277" rel="#L277">277</span>
<span id="L278" rel="#L278">278</span>
<span id="L279" rel="#L279">279</span>
<span id="L280" rel="#L280">280</span>
<span id="L281" rel="#L281">281</span>
<span id="L282" rel="#L282">282</span>
<span id="L283" rel="#L283">283</span>
<span id="L284" rel="#L284">284</span>
<span id="L285" rel="#L285">285</span>
<span id="L286" rel="#L286">286</span>
<span id="L287" rel="#L287">287</span>
<span id="L288" rel="#L288">288</span>
<span id="L289" rel="#L289">289</span>
<span id="L290" rel="#L290">290</span>
<span id="L291" rel="#L291">291</span>
<span id="L292" rel="#L292">292</span>
<span id="L293" rel="#L293">293</span>
<span id="L294" rel="#L294">294</span>
<span id="L295" rel="#L295">295</span>
<span id="L296" rel="#L296">296</span>
<span id="L297" rel="#L297">297</span>
<span id="L298" rel="#L298">298</span>
<span id="L299" rel="#L299">299</span>
<span id="L300" rel="#L300">300</span>
<span id="L301" rel="#L301">301</span>
<span id="L302" rel="#L302">302</span>
<span id="L303" rel="#L303">303</span>
<span id="L304" rel="#L304">304</span>
<span id="L305" rel="#L305">305</span>
<span id="L306" rel="#L306">306</span>
<span id="L307" rel="#L307">307</span>
<span id="L308" rel="#L308">308</span>
<span id="L309" rel="#L309">309</span>
<span id="L310" rel="#L310">310</span>
<span id="L311" rel="#L311">311</span>
<span id="L312" rel="#L312">312</span>
<span id="L313" rel="#L313">313</span>
<span id="L314" rel="#L314">314</span>
<span id="L315" rel="#L315">315</span>
<span id="L316" rel="#L316">316</span>
<span id="L317" rel="#L317">317</span>
<span id="L318" rel="#L318">318</span>
<span id="L319" rel="#L319">319</span>
<span id="L320" rel="#L320">320</span>
<span id="L321" rel="#L321">321</span>
<span id="L322" rel="#L322">322</span>
<span id="L323" rel="#L323">323</span>
<span id="L324" rel="#L324">324</span>
<span id="L325" rel="#L325">325</span>
<span id="L326" rel="#L326">326</span>
<span id="L327" rel="#L327">327</span>
<span id="L328" rel="#L328">328</span>
<span id="L329" rel="#L329">329</span>
<span id="L330" rel="#L330">330</span>
<span id="L331" rel="#L331">331</span>
<span id="L332" rel="#L332">332</span>
<span id="L333" rel="#L333">333</span>
<span id="L334" rel="#L334">334</span>
<span id="L335" rel="#L335">335</span>
<span id="L336" rel="#L336">336</span>
<span id="L337" rel="#L337">337</span>
<span id="L338" rel="#L338">338</span>
<span id="L339" rel="#L339">339</span>
<span id="L340" rel="#L340">340</span>
<span id="L341" rel="#L341">341</span>
<span id="L342" rel="#L342">342</span>
<span id="L343" rel="#L343">343</span>
<span id="L344" rel="#L344">344</span>
<span id="L345" rel="#L345">345</span>
<span id="L346" rel="#L346">346</span>
<span id="L347" rel="#L347">347</span>
<span id="L348" rel="#L348">348</span>
<span id="L349" rel="#L349">349</span>
<span id="L350" rel="#L350">350</span>
<span id="L351" rel="#L351">351</span>
<span id="L352" rel="#L352">352</span>
<span id="L353" rel="#L353">353</span>
<span id="L354" rel="#L354">354</span>
<span id="L355" rel="#L355">355</span>
<span id="L356" rel="#L356">356</span>
<span id="L357" rel="#L357">357</span>
<span id="L358" rel="#L358">358</span>
<span id="L359" rel="#L359">359</span>
<span id="L360" rel="#L360">360</span>
<span id="L361" rel="#L361">361</span>
<span id="L362" rel="#L362">362</span>
<span id="L363" rel="#L363">363</span>
<span id="L364" rel="#L364">364</span>
<span id="L365" rel="#L365">365</span>
<span id="L366" rel="#L366">366</span>
<span id="L367" rel="#L367">367</span>
<span id="L368" rel="#L368">368</span>
<span id="L369" rel="#L369">369</span>
<span id="L370" rel="#L370">370</span>
<span id="L371" rel="#L371">371</span>
<span id="L372" rel="#L372">372</span>
<span id="L373" rel="#L373">373</span>
<span id="L374" rel="#L374">374</span>
<span id="L375" rel="#L375">375</span>
<span id="L376" rel="#L376">376</span>
<span id="L377" rel="#L377">377</span>
<span id="L378" rel="#L378">378</span>
<span id="L379" rel="#L379">379</span>
<span id="L380" rel="#L380">380</span>
<span id="L381" rel="#L381">381</span>
<span id="L382" rel="#L382">382</span>
<span id="L383" rel="#L383">383</span>
<span id="L384" rel="#L384">384</span>
<span id="L385" rel="#L385">385</span>
<span id="L386" rel="#L386">386</span>
<span id="L387" rel="#L387">387</span>
<span id="L388" rel="#L388">388</span>
<span id="L389" rel="#L389">389</span>
<span id="L390" rel="#L390">390</span>
<span id="L391" rel="#L391">391</span>
<span id="L392" rel="#L392">392</span>
<span id="L393" rel="#L393">393</span>
<span id="L394" rel="#L394">394</span>
<span id="L395" rel="#L395">395</span>
<span id="L396" rel="#L396">396</span>
<span id="L397" rel="#L397">397</span>
<span id="L398" rel="#L398">398</span>
<span id="L399" rel="#L399">399</span>
<span id="L400" rel="#L400">400</span>
<span id="L401" rel="#L401">401</span>
<span id="L402" rel="#L402">402</span>
<span id="L403" rel="#L403">403</span>
<span id="L404" rel="#L404">404</span>
<span id="L405" rel="#L405">405</span>
<span id="L406" rel="#L406">406</span>
<span id="L407" rel="#L407">407</span>
<span id="L408" rel="#L408">408</span>
<span id="L409" rel="#L409">409</span>
<span id="L410" rel="#L410">410</span>
<span id="L411" rel="#L411">411</span>
<span id="L412" rel="#L412">412</span>
<span id="L413" rel="#L413">413</span>
<span id="L414" rel="#L414">414</span>
<span id="L415" rel="#L415">415</span>
<span id="L416" rel="#L416">416</span>
<span id="L417" rel="#L417">417</span>
<span id="L418" rel="#L418">418</span>
<span id="L419" rel="#L419">419</span>
<span id="L420" rel="#L420">420</span>
<span id="L421" rel="#L421">421</span>
<span id="L422" rel="#L422">422</span>
<span id="L423" rel="#L423">423</span>
<span id="L424" rel="#L424">424</span>
<span id="L425" rel="#L425">425</span>
<span id="L426" rel="#L426">426</span>
<span id="L427" rel="#L427">427</span>
<span id="L428" rel="#L428">428</span>
<span id="L429" rel="#L429">429</span>
<span id="L430" rel="#L430">430</span>
<span id="L431" rel="#L431">431</span>
<span id="L432" rel="#L432">432</span>
<span id="L433" rel="#L433">433</span>
<span id="L434" rel="#L434">434</span>
<span id="L435" rel="#L435">435</span>
<span id="L436" rel="#L436">436</span>
<span id="L437" rel="#L437">437</span>
<span id="L438" rel="#L438">438</span>
<span id="L439" rel="#L439">439</span>
<span id="L440" rel="#L440">440</span>
<span id="L441" rel="#L441">441</span>
<span id="L442" rel="#L442">442</span>
<span id="L443" rel="#L443">443</span>
<span id="L444" rel="#L444">444</span>
<span id="L445" rel="#L445">445</span>
<span id="L446" rel="#L446">446</span>
<span id="L447" rel="#L447">447</span>
<span id="L448" rel="#L448">448</span>
<span id="L449" rel="#L449">449</span>
<span id="L450" rel="#L450">450</span>
<span id="L451" rel="#L451">451</span>
<span id="L452" rel="#L452">452</span>
<span id="L453" rel="#L453">453</span>
<span id="L454" rel="#L454">454</span>
<span id="L455" rel="#L455">455</span>
<span id="L456" rel="#L456">456</span>
<span id="L457" rel="#L457">457</span>
<span id="L458" rel="#L458">458</span>
<span id="L459" rel="#L459">459</span>
<span id="L460" rel="#L460">460</span>
<span id="L461" rel="#L461">461</span>
<span id="L462" rel="#L462">462</span>
<span id="L463" rel="#L463">463</span>
<span id="L464" rel="#L464">464</span>
</pre>
          </td>
          <td width="100%">
                <div class="highlight"><pre><div class='line' id='LC1'><span class="c">% </span></div><div class='line' id='LC2'><span class="c">% Annual Cognitive Science Conference Format</span></div><div class='line' id='LC3'><br/></div><div class='line' id='LC4'><br/></div><div class='line' id='LC5'><span class="k">\documentclass</span><span class="na">[10pt,letterpaper]</span><span class="nb">{</span>article<span class="nb">}</span></div><div class='line' id='LC6'><br/></div><div class='line' id='LC7'><span class="k">\usepackage</span><span class="nb">{</span>mathtools<span class="nb">}</span></div><div class='line' id='LC8'><span class="k">\usepackage</span><span class="nb">{</span>cogsci<span class="nb">}</span></div><div class='line' id='LC9'><span class="k">\usepackage</span><span class="nb">{</span>pslatex<span class="nb">}</span></div><div class='line' id='LC10'><span class="k">\usepackage</span><span class="nb">{</span>apacite<span class="nb">}</span></div><div class='line' id='LC11'><br/></div><div class='line' id='LC12'><br/></div><div class='line' id='LC13'><span class="k">\title</span><span class="nb">{</span>A Bayesian Framework for Learning Regular Expressions<span class="nb">}</span></div><div class='line' id='LC14'>&nbsp;</div><div class='line' id='LC15'><span class="k">\author</span><span class="nb">{{</span><span class="k">\large</span> <span class="k">\bf</span><span class="nb">{</span>P. Thomas Barthelemy<span class="nb">}</span> (bartho@stanford.edu)<span class="nb">}</span> <span class="k">\\</span></div><div class='line' id='LC16'>&nbsp;&nbsp;Department of Computer Science, 353 Serra Mall <span class="k">\\</span></div><div class='line' id='LC17'>&nbsp;&nbsp;Stanford, CA 94305 USA</div><div class='line' id='LC18'>&nbsp;&nbsp;<span class="k">\AND</span> <span class="nb">{</span><span class="k">\large</span> <span class="k">\bf</span> Nicholas Borg (nickborg@stanford.edu)<span class="nb">}</span> <span class="k">\\</span></div><div class='line' id='LC19'>&nbsp;&nbsp;Symbolic Systems Program, Margaret Jacks Hall <span class="k">\\</span></div><div class='line' id='LC20'>&nbsp;&nbsp;Stanford, CA 94305 USA<span class="nb">}</span></div><div class='line' id='LC21'><br/></div><div class='line' id='LC22'><br/></div><div class='line' id='LC23'><span class="k">\begin</span><span class="nb">{</span>document<span class="nb">}</span></div><div class='line' id='LC24'><br/></div><div class='line' id='LC25'><span class="k">\maketitle</span></div><div class='line' id='LC26'><br/></div><div class='line' id='LC27'><br/></div><div class='line' id='LC28'><span class="k">\begin</span><span class="nb">{</span>abstract<span class="nb">}</span></div><div class='line' id='LC29'>Regular expressions are used to define a set of strings that match a certain pattern. In the following experiment, we posit that the means by which humans judge similarity of words can be represented by a method of inference using regular expressions. In short, there is a tradeoff of generality and specificity that allows for a selection of an optimal regular expression to categorize a list of input strings.</div><div class='line' id='LC30'><br/></div><div class='line' id='LC31'><span class="k">\textbf</span><span class="nb">{</span>Keywords:<span class="nb">}</span> </div><div class='line' id='LC32'>Model inference, regular language, regular expression, deterministic finite automaton.</div><div class='line' id='LC33'><span class="k">\end</span><span class="nb">{</span>abstract<span class="nb">}</span></div><div class='line' id='LC34'><br/></div><div class='line' id='LC35'><br/></div><div class='line' id='LC36'><span class="k">\section</span><span class="nb">{</span>Overview and Motivation<span class="nb">}</span></div><div class='line' id='LC37'>Understanding the human ability to find structure in data is integral to both cognitive psychology and artificial intelligence. In recent years, a number of studies have shown that human inference across a number of tasks may be Bayesian at some level of analysis. Inspired by the treatment of the number game in <span class="k">\citeA</span><span class="nb">{</span>tenenbaum1999bayesian<span class="nb">}</span>, we propose a regular expression game as follows: Given a set <span class="s">$</span><span class="nb">S</span><span class="s">$</span> of input strings drawn from a language <span class="s">$</span><span class="nv">\mathcal</span><span class="nb">{L}</span><span class="s">$</span>, what is the likelihood that a new string <span class="s">$</span><span class="nb">s</span><span class="s">$</span> is in <span class="s">$</span><span class="nv">\mathcal</span><span class="nb">{L}</span><span class="s">$</span>. Playing such a game requires inferring the rules that govern <span class="s">$</span><span class="nv">\mathcal</span><span class="nb">{L}</span><span class="s">$</span> from <span class="s">$</span><span class="nb">S</span><span class="s">$</span>. Noting that humans perform this task  with success on a daily basis in writing regular expressions to process string data, we considered the hypothesis that human inference of regular expressions is at some level Bayesian. </div><div class='line' id='LC38'><br/></div><div class='line' id='LC39'>In doing so, we first generated a Bayesian model that, given a set of input strings <span class="s">$</span><span class="nb">S</span><span class="s">$</span>, generated a regular expression that describes <span class="s">$</span><span class="nb">S</span><span class="s">$</span> with the highest posterior probability. This model is discussed in the first part of this paper. We then ran a behavioral experiment in which subjects were asked whether or not a given string <span class="s">$</span><span class="nb">s</span><span class="s">$</span> fits with a set of strings <span class="s">$</span><span class="nb">S</span><span class="s">$</span>. We compare the result of that experiment with results generated given an extension of our model.  </div><div class='line' id='LC40'><br/></div><div class='line' id='LC41'><span class="k">\section</span><span class="nb">{</span>Description of the Model<span class="nb">}</span></div><div class='line' id='LC42'><br/></div><div class='line' id='LC43'>Our model has two parts: We generate a set of regular expressions which describe a set of input stings <span class="s">$</span><span class="nb">S</span><span class="s">$</span>, and we find the regular expression in that set with the highest posterior probability. The posterior probability is proportional to the product of the hypothesis&#39; prior probability and its likelihood.</div><div class='line' id='LC44'><br/></div><div class='line' id='LC45'>Below, we define the prior probability such that our model biases smaller hypotheses, and the likelihood of generating a given hypothesis is inversely proportional to its cost. Our model aims to balance generality with specificity. For example, given <span class="k">\verb</span>!abc, abd!, we wish to favor the regex <span class="k">\verb</span>!ab(c|d)! over <span class="k">\verb</span>!A*! where <span class="k">\verb</span>!A! represents any letter,<span class="k">\footnote</span><span class="nb">{</span>See the Appendix for a brief summary of our regex notation.<span class="nb">}</span> which is less likely given the size principle. At the same time, given <span class="k">\verb</span>!aba!, <span class="k">\verb</span>!abc!, <span class="k">\verb</span>!abd!, <span class="k">\verb</span>!abe!, <span class="k">\verb</span>!abf!, we may wish to favor <span class="k">\verb</span>!abA! over <span class="k">\verb</span>!ab(a|c|d|e|f)!.</div><div class='line' id='LC46'><br/></div><div class='line' id='LC47'><span class="k">\subsection</span><span class="nb">{</span>Representation of Hypotheses<span class="nb">}</span></div><div class='line' id='LC48'><br/></div><div class='line' id='LC49'>Regular expressions are equivalent to deterministic finite automata (DFA). Because DFA are often easier to reason with, we use them to represent regular expressions in our model (See Figure~<span class="k">\ref</span><span class="nb">{</span>base<span class="nb">_</span>hypothesis<span class="nb">}</span> for an example DFA.)</div><div class='line' id='LC50'><br/></div><div class='line' id='LC51'>Further, we define the probability with which a DFA would stop generating a string at a start state . For instance, given the regular expression <span class="k">\verb</span>!a*!, we would need to define the probability of generating strings <span class="k">\verb</span>!a!, <span class="k">\verb</span>!aa!, <span class="k">\verb</span>!aaa!, etc.</div><div class='line' id='LC52'><br/></div><div class='line' id='LC53'>Our generative version of the DFA is not to be confused with a probabilistic atumaton, which has transition probabilities much as described above and a final acceptance threshold. Our generative automata does not have this threshold, and instead has an added parameter which controls the likelihood of stopping to generate a string.</div><div class='line' id='LC54'><br/></div><div class='line' id='LC55'>As our alphabet, we use the set of all lowercase letters. This set, we feel, has an intuitive boundary, as subjects seeing strings of all lowercase letters are unlikely to suppose an arbitrary uppercase letter or special character, and thus we assume the expressiveness of wildcards to be fixed.</div><div class='line' id='LC56'><br/></div><div class='line' id='LC57'><span class="c">% TODO: explain character scope!</span></div><div class='line' id='LC58'><br/></div><div class='line' id='LC59'><span class="k">\subsection</span><span class="nb">{</span>Generating Hypotheses<span class="nb">}</span></div><div class='line' id='LC60'>The hypothesis space is initiated with the base hypothesis, the term applied to a DFA that accepts strictly the input strings. We begin generating a series of consecutive states, where the transition between the first state and the second state is governed by the first letter of the first string, the transition between the second and the third is governend by the the second letter, and so on. For each additional string, if the first i characters overlap with those of any of the first i characters of one previous string but differ in the <span class="s">$</span><span class="nb">i</span><span class="o">+</span><span class="m">1</span><span class="s">$</span>th, we add no new states to the DFA until the transition between the i-1th state and the ith state. We then generate new states for the remaining characters of the string. See Figure~<span class="k">\ref</span><span class="nb">{</span>base<span class="nb">_</span>hypothesis<span class="nb">}</span> for an example of the first hypothesis generated from a set <span class="s">$</span><span class="nb">S </span><span class="o">=</span><span class="nb"> </span><span class="s">$</span><span class="k">\{\verb</span>!ababb!, <span class="k">\verb</span>!abbab!<span class="k">\}</span> .</div><div class='line' id='LC61'>&nbsp;&nbsp;<span class="k">\begin</span><span class="nb">{</span>figure<span class="nb">}</span>[ht]</div><div class='line' id='LC62'><span class="k">\begin</span><span class="nb">{</span>center<span class="nb">}</span></div><div class='line' id='LC63'><span class="k">\begin</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC64'><span class="k">\includegraphics</span><span class="na">[scale=0.4]</span><span class="nb">{</span>readin.png<span class="nb">}</span></div><div class='line' id='LC65'><span class="k">\end</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC66'><span class="k">\end</span><span class="nb">{</span>center<span class="nb">}</span></div><div class='line' id='LC67'><span class="k">\caption</span><span class="nb">{</span>An example base hypothesis DFA generated from the strings `ababb&#39; and `abbab&#39;. Note that the two strings share the first two states, but that the fifth and tenth are distinct.<span class="nb">}</span> </div><div class='line' id='LC68'><span class="k">\label</span><span class="nb">{</span>base<span class="nb">_</span>hypothesis<span class="nb">}</span></div><div class='line' id='LC69'><span class="k">\end</span><span class="nb">{</span>figure<span class="nb">}</span></div><div class='line' id='LC70'><br/></div><div class='line' id='LC71'>Next, hypotheses are added to the hypothesis space through two processes of generalization: merging and applying wildcards.</div><div class='line' id='LC72'>&nbsp;</div><div class='line' id='LC73'><br/></div><div class='line' id='LC74'><span class="k">\subsubsection</span><span class="nb">{</span>Merging<span class="nb">}</span> Merging is the process of combining states of a DFA. The newly create state has all incoming and outgoing transitions of its predecessors. Thus, the DFA will continue to accept the same set of strings as the original DFA. However, the DFA will often accept more strings after merging, thus merging states results in a more generalized DFA.</div><div class='line' id='LC75'><br/></div><div class='line' id='LC76'>If merging is continued, the DFA will eventually have only one state with all transitions leading to itself. Such a state accepts all strings containing only zero or more of the original characters. A series of merges are illustrated in Figure~<span class="k">\ref</span><span class="nb">{</span>merge-figure<span class="nb">}</span>.</div><div class='line' id='LC77'><br/></div><div class='line' id='LC78'><span class="k">\begin</span><span class="nb">{</span>figure<span class="nb">}</span>[ht]</div><div class='line' id='LC79'><span class="k">\begin</span><span class="nb">{</span>center<span class="nb">}</span></div><div class='line' id='LC80'><span class="k">\begin</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC81'><span class="k">\text</span><span class="nb">{</span>(a)<span class="nb">}</span><span class="k">\includegraphics</span><span class="na">[scale=0.4]</span><span class="nb">{</span>merge-0.png<span class="nb">}</span></div><div class='line' id='LC82'><span class="nb">&amp;</span><span class="k">\text</span><span class="nb">{</span>(b)<span class="nb">}</span><span class="k">\includegraphics</span><span class="na">[scale=0.4]</span><span class="nb">{</span>merge-1.png<span class="nb">}</span></div><div class='line' id='LC83'><span class="nb">&amp;</span><span class="k">\text</span><span class="nb">{</span>(c)<span class="nb">}</span><span class="k">\includegraphics</span><span class="na">[scale=0.4]</span><span class="nb">{</span>merge-2.png<span class="nb">}</span></div><div class='line' id='LC84'><span class="k">\end</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC85'><span class="k">\end</span><span class="nb">{</span>center<span class="nb">}</span></div><div class='line' id='LC86'><span class="k">\caption</span><span class="nb">{</span>An example of state merging. The DFA starts at state 0 and accepts only at the shaded states. The base hypothesis accepts ``b&#39;&#39; and ``ab&#39;&#39; (a). After merging states 0 and 2, the next DFA accepts an infinite set of strings (b). Finally, the merge of states 0 and 1 forces the merge of state 0 and 3 to maintain one unique outbound transition for any state and character (c).<span class="nb">}</span> </div><div class='line' id='LC87'><span class="k">\label</span><span class="nb">{</span>merge-figure<span class="nb">}</span></div><div class='line' id='LC88'><span class="k">\end</span><span class="nb">{</span>figure<span class="nb">}</span></div><div class='line' id='LC89'><br/></div><div class='line' id='LC90'><span class="c">% TODO: mention anti-unification</span></div><div class='line' id='LC91'><br/></div><div class='line' id='LC92'><span class="k">\subsubsection</span><span class="nb">{</span>Wildcards<span class="nb">}</span> Wildcards generalize the DFA by extending state transitions to apply to a class of characters. Figure~<span class="k">\ref</span><span class="nb">{</span>wildcard-figure<span class="nb">}</span> illustrates an example of this.</div><div class='line' id='LC93'><br/></div><div class='line' id='LC94'><span class="k">\begin</span><span class="nb">{</span>figure<span class="nb">}</span>[ht]</div><div class='line' id='LC95'><span class="k">\begin</span><span class="nb">{</span>center<span class="nb">}</span></div><div class='line' id='LC96'><span class="k">\begin</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC97'><span class="k">\text</span><span class="nb">{</span>(a)<span class="nb">}</span><span class="k">\includegraphics</span><span class="na">[scale=0.4]</span><span class="nb">{</span>wildcard-0.png<span class="nb">}</span></div><div class='line' id='LC98'><span class="nb">&amp;</span><span class="k">\text</span><span class="nb">{</span>(b)<span class="nb">}</span><span class="k">\includegraphics</span><span class="na">[scale=0.4]</span><span class="nb">{</span>wildcard-1.png<span class="nb">}</span></div><div class='line' id='LC99'><span class="k">\end</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC100'><span class="k">\end</span><span class="nb">{</span>center<span class="nb">}</span></div><div class='line' id='LC101'><span class="k">\caption</span><span class="nb">{</span>An example of inserting a wildcard, ``A&#39;&#39;, which represents any character. Replacing a wildcard for one of the outbound transitions of state 0 has forced a merge of states 1 and 3.<span class="nb">}</span> </div><div class='line' id='LC102'><span class="k">\label</span><span class="nb">{</span>wildcard-figure<span class="nb">}</span></div><div class='line' id='LC103'><span class="k">\end</span><span class="nb">{</span>figure<span class="nb">}</span></div><div class='line' id='LC104'><br/></div><div class='line' id='LC105'><span class="k">\subsubsection</span><span class="nb">{</span>Enforcing Determinism<span class="nb">}</span> Merging states and replacing transitions with wildcards can allow one state to have non-unique outbound transitions. However, this turns our representation into a nondeterministic finite automaton (NFA) instead of a DFA. Both are equivalent, but the latter afford easier manipulation for our experiment. Thus, to convert the NFA back to a DFA, we use the subset construction, the effect of which is illustrated in both Figure~<span class="k">\ref</span><span class="nb">{</span>merge-figure<span class="nb">}</span> and Figure~<span class="k">\ref</span><span class="nb">{</span>wildcard-figure<span class="nb">}</span>.</div><div class='line' id='LC106'><br/></div><div class='line' id='LC107'><span class="k">\subsection</span><span class="nb">{</span>Searching the Hypothesis Space<span class="nb">}</span> The number of hypotheses grows quickly with the number of states in the base hypothesis. As a lower bound, the number of hypotheses is at least as large as the number of possible starting nodes. Because the starting node could be merged with any subset of the remaining <span class="s">$</span><span class="nb">n</span><span class="o">-</span><span class="m">1</span><span class="s">$</span> nodes in the DFA, there are <span class="s">$</span><span class="m">2</span><span class="nb">^{n</span><span class="o">-</span><span class="m">1</span><span class="nb">}</span><span class="o">=</span><span class="nv">\Omega</span><span class="o">(</span><span class="m">2</span><span class="nb">^n</span><span class="o">)</span><span class="s">$</span> possible start states. Thus, the hypothesis space is at least exponential in the number of starting nodes in the regular expression.</div><div class='line' id='LC108'><br/></div><div class='line' id='LC109'>Similarly, the number of possible regex variants created by simply replacing transitions with wildcards is given by <span class="s">$</span><span class="m">2</span><span class="nb">^m</span><span class="s">$</span> where <span class="s">$</span><span class="nb">m</span><span class="s">$</span> is the number of transitions. Thus, simply modifying expressions to include wildcards is also exponential.</div><div class='line' id='LC110'><br/></div><div class='line' id='LC111'>An upper bound for the complexity of model merging is much more extreme. Performing subset construction for an NFA with <span class="s">$</span><span class="nb">n</span><span class="s">$</span> states can produce, in the worst case, a DFA with <span class="s">$</span><span class="m">2</span><span class="nb">^n</span><span class="s">$</span> states. Assuming that any subset of these states also produces a valid DFA, we can have at most <span class="s">$</span><span class="m">2</span><span class="nb">^{</span><span class="m">2</span><span class="nb">^n</span><span class="o">-</span><span class="m">1</span><span class="nb">}</span><span class="s">$</span> DFA.</div><div class='line' id='LC112'><span class="c">% \footnote{See the Appendix for detailed sizing of the hypothesis space.} Actually, this seems inaccurate.</span></div><div class='line' id='LC113'><br/></div><div class='line' id='LC114'>We resolve this issue by performing beam search. In beam first, we perform breadth first search, but for any given depth of the search tree, we only consider children of the <span class="s">$</span><span class="nb">m</span><span class="s">$</span> most promising hypotheses. In our case, the most promising hypotheses were defined as those with the highest posterior probability. (For a more detailed discussion of Beam Search, see <span class="k">\cite</span><span class="nb">{</span>HwangStuhlmullerGoodman2011<span class="nb">}</span></div><div class='line' id='LC115'><br/></div><div class='line' id='LC116'><br/></div><div class='line' id='LC117'><span class="k">\section</span><span class="nb">{</span>Inference with DFA<span class="nb">}</span></div><div class='line' id='LC118'>The posterior probability for a DFA as a hypothesis is given by the following.</div><div class='line' id='LC119'><span class="k">\begin</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC120'>	p(h|S) <span class="k">\propto</span> p(S|h)p(h)</div><div class='line' id='LC121'><span class="k">\end</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC122'>Thus, inferring the best DFA given a set of input strings was essentially a problem in balancing simplicity with conservation of belief. That is, DFA were biased a priori for being simpler, but simpler DFA tended to accept more strings.</div><div class='line' id='LC123'><br/></div><div class='line' id='LC124'><span class="k">\subsection</span><span class="nb">{</span>Prior Probability<span class="nb">}</span></div><div class='line' id='LC125'>We assigned the prior probability to the hypotheses (i.e. regexes) based on the size of the DFA and the number of transitions. In the following equation, <span class="s">$</span><span class="nb">|S|</span><span class="s">$</span> is the number of states in the DFA and <span class="s">$</span><span class="nb">|</span><span class="nv">\delta</span><span class="nb">|</span><span class="s">$</span> is the number of tranisitons in the DFA, where a wildcard counts as only one transition. </div><div class='line' id='LC126'><span class="k">\begin</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC127'>	P(h) <span class="k">\propto</span> e<span class="nb">^{</span>-<span class="k">\alpha</span>|S| - <span class="k">\beta</span>|<span class="k">\delta</span>|<span class="nb">}</span></div><div class='line' id='LC128'><span class="k">\end</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC129'>Thus, we have biased smaller DFA that use wildcards instead of literal characters. [cite Program Merging model]</div><div class='line' id='LC130'><br/></div><div class='line' id='LC131'><span class="k">\subsection</span><span class="nb">{</span>Likelihood<span class="nb">}</span></div><div class='line' id='LC132'>To define the likelihood of producing certain strings, we turn a DFA into a generative automaton with a probability distribution over generated strings. Letting the input strings be independent, the probability of generating the set of strings <span class="s">$</span><span class="nb">S</span><span class="s">$</span> is equal to the product of the probabilities of generating each string.</div><div class='line' id='LC133'><span class="k">\begin</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC134'>	p(S) = <span class="k">\prod</span><span class="nb">_{</span>s <span class="k">\in</span> S<span class="nb">}</span> p(s)</div><div class='line' id='LC135'><span class="k">\end</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC136'><br/></div><div class='line' id='LC137'>Let the string <span class="s">$</span><span class="nb">s</span><span class="s">$</span> be the concatenation of characters <span class="s">$</span><span class="nv">\{</span><span class="nb">a_</span><span class="m">0</span><span class="nb">,a_</span><span class="m">1</span><span class="nb">,...,a_{n</span><span class="o">-</span><span class="m">1</span><span class="nb">}</span><span class="nv">\}</span><span class="s">$</span>. We define the probability of generating a string as the product of the probability of generating the letters in sequence <span class="s">$</span><span class="nb">p</span><span class="o">(</span><span class="nb">a_</span><span class="m">0</span><span class="nb">a_</span><span class="m">1</span><span class="nb">...a_{n</span><span class="o">-</span><span class="m">1</span><span class="nb">}</span><span class="o">)</span><span class="s">$</span> and the product of stopping the string generation on an accept state <span class="s">$</span><span class="nb">p</span><span class="o">(</span><span class="nv">\text</span><span class="nb">{stop}</span><span class="o">)</span><span class="s">$</span>.</div><div class='line' id='LC138'><span class="k">\begin</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC139'>	p(s) = p(a<span class="nb">_</span>1a<span class="nb">_</span>2...a<span class="nb">_</span>n)p(<span class="k">\text</span><span class="nb">{</span>stop<span class="nb">}</span>)</div><div class='line' id='LC140'><span class="k">\end</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC141'><br/></div><div class='line' id='LC142'>We define the probability of generating the characters as the product of the probabilities of making each state transition.</div><div class='line' id='LC143'><span class="k">\begin</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC144'>	p(a<span class="nb">_</span>1a<span class="nb">_</span>2 ... a<span class="nb">_</span>n) = <span class="k">\prod</span><span class="nb">_{</span>i=0<span class="nb">}^{</span>n-1<span class="nb">}</span>p(<span class="k">\delta</span>(r<span class="nb">_</span>i, a<span class="nb">_</span>i))</div><div class='line' id='LC145'><span class="k">\end</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC146'>where <span class="s">$</span><span class="nb">r_i</span><span class="s">$</span> is the state reached after the first <span class="s">$</span><span class="nb">i</span><span class="s">$</span> characters, and <span class="s">$</span><span class="nb">r_</span><span class="m">0</span><span class="s">$</span> is the start state.</div><div class='line' id='LC147'><br/></div><div class='line' id='LC148'>To simplify inference, we assume uniform probability over all possible outbound transitions for a given state. Formally, the probability of making a state transition is given by:</div><div class='line' id='LC149'><span class="k">\begin</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC150'>	p(<span class="k">\delta</span>(r<span class="nb">_</span>i, a<span class="nb">_{</span>i+1<span class="nb">}</span>)) <span class="nb">&amp;</span>= <span class="k">\frac</span><span class="nb">{</span>1<span class="nb">}{</span><span class="k">\sum</span><span class="nb">_{</span>a<span class="k">\in</span> A<span class="nb">}</span> 1<span class="k">\{\delta</span>(r<span class="nb">_</span>i, a)<span class="k">\in</span> Q<span class="k">\}</span><span class="nb">}</span></div><div class='line' id='LC151'><span class="k">\end</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC152'>where <span class="s">$</span><span class="m">1</span><span class="nv">\{\cdot\}</span><span class="s">$</span> is the indicator function, which evaluates to <span class="s">$</span><span class="m">1</span><span class="s">$</span> if the enclosed proposition is true and <span class="s">$</span><span class="m">0</span><span class="s">$</span> otherwise. <span class="s">$</span><span class="nv">\delta</span><span class="o">(</span><span class="nb">r_i, a</span><span class="o">)</span><span class="nv">\in</span><span class="nb"> Q</span><span class="s">$</span> indicates that the state transition leads to another state of the DFA, so the denominator counts the number of characters for which a transition is defined. In the case of the wildcard <span class="k">\verb</span>!A!, the transition is defined for all 26 letters, so the denominator is 26.</div><div class='line' id='LC153'><br/></div><div class='line' id='LC154'>For the cases in which a string may alternatively be accepted or continue, one must assign a probability for accepting or continuing the string. For instance, consider the DFA in part (c) of Figure~<span class="k">\ref</span><span class="nb">{</span>merge-figure<span class="nb">}</span>. The DFA may generate the empty string, <span class="k">\verb</span>!E!, or it may generate any arbitrary string with <span class="k">\verb</span>!a! and <span class="k">\verb</span>!b!, like <span class="k">\verb</span>!ababbba!. We define the probability of stopping at a given string to be a fixed value <span class="s">$</span><span class="nv">\tau</span><span class="s">$</span>. This makes shorter strings more likely to be generated.</div><div class='line' id='LC155'><br/></div><div class='line' id='LC156'>Instead of leaving <span class="s">$</span><span class="nv">\tau</span><span class="s">$</span> as a parameter of the model, we assume a uniform prior over the parameter and marginalize it when calculating the likelihood of a string. So, let <span class="s">$</span><span class="nb">b</span><span class="s">$</span> be the number of times the DFA skipped an accept state before accepting. That is, <span class="s">$</span><span class="nb">b</span><span class="s">$</span> is given by the equation below, where <span class="s">$</span><span class="nb">F</span><span class="s">$</span> is the set of accept states.</div><div class='line' id='LC157'><span class="k">\begin</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC158'>	b = <span class="k">\sum</span><span class="nb">_{</span>i=1<span class="nb">}^{</span>n-1<span class="nb">}</span> 1 <span class="k">\{</span> r<span class="nb">_</span>i <span class="k">\in</span> F<span class="k">\}</span></div><div class='line' id='LC159'><span class="k">\end</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC160'><br/></div><div class='line' id='LC161'><span class="c">% TODO: Cite this</span></div><div class='line' id='LC162'><br/></div><div class='line' id='LC163'>The probability of stopping at an accept state is given by the equation below. The condition <span class="s">$</span><span class="nb">r_n </span><span class="nv">\in</span><span class="nb"> F </span><span class="nv">\wedge</span><span class="nb"> </span><span class="nv">\exists</span><span class="nb"> a : </span><span class="nv">\delta</span><span class="o">(</span><span class="nb">r_n, a</span><span class="o">)</span><span class="nv">\in</span><span class="nb"> Q</span><span class="s">$</span> states that the last state reached by the string is an accept state, and that that state has at least one outbound transition.</div><div class='line' id='LC164'><span class="k">\begin</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC165'>	p(<span class="k">\text</span><span class="nb">{</span>stop<span class="nb">}</span>) <span class="nb">&amp;</span>= <span class="k">\left\{</span></div><div class='line' id='LC166'>		<span class="k">\begin</span><span class="nb">{</span>array<span class="nb">}{</span>rl<span class="nb">}</span></div><div class='line' id='LC167'>		<span class="k">\tau</span> (1-<span class="k">\tau</span>)<span class="nb">^</span>b <span class="nb">&amp;</span> r<span class="nb">_</span>n <span class="k">\in</span> F <span class="k">\wedge</span> <span class="k">\exists</span> a : <span class="k">\delta</span>(r<span class="nb">_</span>n, a)<span class="k">\in</span> Q</div><div class='line' id='LC168'>		<span class="k">\\</span> (1-<span class="k">\tau</span>)<span class="nb">^</span>b <span class="nb">&amp;</span> r<span class="nb">_</span>n <span class="k">\in</span> F <span class="k">\wedge</span> <span class="k">\not</span> <span class="k">\exists</span> a : <span class="k">\delta</span>(r<span class="nb">_</span>n, a)<span class="k">\in</span> Q</div><div class='line' id='LC169'>		<span class="k">\\</span> 0 <span class="nb">&amp;</span> <span class="k">\text</span><span class="nb">{</span>otherwise<span class="nb">}</span></div><div class='line' id='LC170'>		<span class="k">\end</span><span class="nb">{</span>array<span class="nb">}</span></div><div class='line' id='LC171'>		<span class="k">\right</span>.</div><div class='line' id='LC172'><span class="k">\end</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC173'><br/></div><div class='line' id='LC174'>Marginalizing over all possible values of <span class="s">$</span><span class="nv">\tau</span><span class="s">$</span>, we have:</div><div class='line' id='LC175'><span class="k">\begin</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC176'>	<span class="k">\int</span><span class="nb">_</span>0<span class="nb">^</span>1 <span class="k">\tau</span> (1-<span class="k">\tau</span>)<span class="nb">^</span>b <span class="nb">&amp;</span>= <span class="k">\frac</span><span class="nb">{</span>1<span class="nb">}{</span>b<span class="nb">^</span>2+3b+1<span class="nb">}</span></div><div class='line' id='LC177'>	<span class="k">\\</span> <span class="k">\int</span><span class="nb">_</span>0<span class="nb">^</span>1 (1-<span class="k">\tau</span>)<span class="nb">^</span>b <span class="nb">&amp;</span>= <span class="k">\frac</span><span class="nb">{</span>1<span class="nb">}{</span>b+1<span class="nb">}</span></div><div class='line' id='LC178'><span class="k">\end</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC179'><br/></div><div class='line' id='LC180'>Our equation for <span class="s">$</span><span class="nb">p</span><span class="o">(</span><span class="nv">\text</span><span class="nb">{stop}</span><span class="o">)</span><span class="s">$</span> is thus:</div><div class='line' id='LC181'><span class="k">\begin</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC182'>	p(<span class="k">\text</span><span class="nb">{</span>stop<span class="nb">}</span>) <span class="nb">&amp;</span>= <span class="k">\left\{</span></div><div class='line' id='LC183'>		<span class="k">\begin</span><span class="nb">{</span>array<span class="nb">}{</span>rl<span class="nb">}</span></div><div class='line' id='LC184'>		<span class="k">\frac</span><span class="nb">{</span>1<span class="nb">}{</span>b<span class="nb">^</span>2+3b+1<span class="nb">}</span> <span class="nb">&amp;</span> r<span class="nb">_</span>n <span class="k">\in</span> F <span class="k">\wedge</span> <span class="k">\exists</span> a : <span class="k">\delta</span>(r<span class="nb">_</span>n, a)</div><div class='line' id='LC185'>		<span class="k">\\</span> <span class="k">\frac</span><span class="nb">{</span>1<span class="nb">}{</span>b+1<span class="nb">}</span> <span class="nb">&amp;</span> r<span class="nb">_</span>n <span class="k">\in</span> F <span class="k">\wedge</span> <span class="k">\not</span> <span class="k">\exists</span> a : <span class="k">\delta</span>(r<span class="nb">_</span>n, a)</div><div class='line' id='LC186'>		<span class="k">\\</span> 0 <span class="nb">&amp;</span> <span class="k">\text</span><span class="nb">{</span>otherwise<span class="nb">}</span></div><div class='line' id='LC187'>		<span class="k">\end</span><span class="nb">{</span>array<span class="nb">}</span></div><div class='line' id='LC188'>		<span class="k">\right</span>.</div><div class='line' id='LC189'><span class="k">\end</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC190'><br/></div><div class='line' id='LC191'><span class="k">\section</span><span class="nb">{</span>Experiment Design and Model Fitting<span class="nb">}</span></div><div class='line' id='LC192'>Having generated the above Bayesian model for selection of the regular expression with the highest posterior probability, we then conducted a behavioral experiment.  41 subjects (Median age 34, std-dev 13.3) were surveyed on Amazon Mechanical Turk. Two groups (one of 20, one of 21) were presented with 10 yes or no questions in which they were asked to determine whether a given string `fit&#39; with a list of strings. A sample question was as follows: <span class="k">\\</span></div><div class='line' id='LC193'><span class="k">\begin</span><span class="nb">{</span>center<span class="nb">}</span>Given the list:</div><div class='line' id='LC194'><br/></div><div class='line' id='LC195'><span class="k">\textbf</span><span class="nb">{</span>yyyt <span class="k">\\</span></div><div class='line' id='LC196'>yt <span class="k">\\</span></div><div class='line' id='LC197'>yyyt <span class="k">\\</span></div><div class='line' id='LC198'>yyyt <span class="k">\\</span> </div><div class='line' id='LC199'>yt<span class="nb">}</span></div><div class='line' id='LC200'><br/></div><div class='line' id='LC201'>Does <span class="k">\textbf</span><span class="nb">{</span>yyt<span class="nb">}</span> fit?</div><div class='line' id='LC202'><span class="k">\end</span><span class="nb">{</span>center<span class="nb">}</span></div><div class='line' id='LC203'><br/></div><div class='line' id='LC204'>The experimental data served two purposes. Using the responses to four out of 23 questions, we set the parameters for alpha and beta. We then used the remaining questions to test our model&#39;s performance in a regex game as shown below. </div><div class='line' id='LC205'><br/></div><div class='line' id='LC206'><span class="k">\subsection</span><span class="nb">{</span>Parameter Fitting<span class="nb">}</span></div><div class='line' id='LC207'>Because the role of <span class="s">$</span><span class="nv">\beta</span><span class="s">$</span> had the effect of biasing models with fewer transitions, it also had a tendency to bias regular expressions with fewer states. For instance, the <span class="s">$</span><span class="nv">\beta</span><span class="s">$</span> term would prefer the regular expression <span class="k">\verb</span>!a*! over <span class="k">\verb</span>!aaaa!. Thus, this term was fitted before fitting the <span class="s">$</span><span class="nv">\alpha</span><span class="s">$</span> term.</div><div class='line' id='LC208'><br/></div><div class='line' id='LC209'>To fit the <span class="s">$</span><span class="nv">\beta</span><span class="s">$</span> term, we observed experiments that would indicate the transition point between a preference for generality over specificity for wildcard-based experiments (e.g. the first two experiments outlined in Table~<span class="k">\ref</span><span class="nb">{</span>parameter-table<span class="nb">}</span>). Utilizing the fact that increasing the ratio of repeated strings to new strings decreases the likelihood of a wildcard, and decreasing the length of the sample strings would increases the likelihood of a wildcard, the first two experiments suggest a decision point. That is, if the DFA model is a valid representation of the subjects&#39; responses, then the <span class="s">$</span><span class="nv">\beta</span><span class="s">$</span> term should be between <span class="s">$</span><span class="m">3</span><span class="s">$</span> and <span class="s">$</span><span class="m">5</span><span class="s">$</span>. For further experimentation, we fixed it at <span class="s">$</span><span class="m">4</span><span class="s">$</span>.</div><div class='line' id='LC210'><br/></div><div class='line' id='LC211'>Fitting the <span class="s">$</span><span class="nv">\alpha</span><span class="s">$</span> term was less clear. In particular, the desired behavior was easily observed with <span class="s">$</span><span class="nv">\alpha</span><span class="o">=</span><span class="m">0</span><span class="nb">,</span><span class="nv">\beta</span><span class="o">=</span><span class="m">4</span><span class="s">$</span>. However, a term of <span class="s">$</span><span class="nv">\alpha</span><span class="o">=</span><span class="m">1</span><span class="s">$</span> was used to bias hypotheses with fewer states given the same number of transitions.</div><div class='line' id='LC212'><br/></div><div class='line' id='LC213'><span class="c">% TODO: move this?</span></div><div class='line' id='LC214'><span class="k">\begin</span><span class="nb">{</span>table<span class="nb">}</span>[!ht]</div><div class='line' id='LC215'><span class="k">\begin</span><span class="nb">{</span>center<span class="nb">}</span> </div><div class='line' id='LC216'><span class="k">\caption</span><span class="nb">{</span>Parameter fitting experiments<span class="nb">}</span> </div><div class='line' id='LC217'><span class="k">\label</span><span class="nb">{</span>parameter-table<span class="nb">}</span> </div><div class='line' id='LC218'><span class="k">\vskip</span> 0.12in</div><div class='line' id='LC219'><span class="k">\begin</span><span class="nb">{</span>tabular<span class="nb">}{</span>llll<span class="nb">}</span> </div><div class='line' id='LC220'><span class="k">\hline</span></div><div class='line' id='LC221'>Input  <span class="nb">&amp;</span> Test <span class="nb">&amp;</span> Subject Accept <span class="nb">&amp;</span> Model Accept</div><div class='line' id='LC222'><span class="c">% \\ \hline</span></div><div class='line' id='LC223'><span class="c">% bhi, bhw, bhi, &amp; bho &amp; 30\% &amp; $\beta &lt; 14$</span></div><div class='line' id='LC224'><span class="c">% \\ bhw, bhw</span></div><div class='line' id='LC225'><span class="k">\\</span> <span class="k">\hline</span></div><div class='line' id='LC226'>efg, hfg, ifg <span class="nb">&amp;</span> ffg <span class="nb">&amp;</span> 45<span class="k">\%</span> <span class="nb">&amp;</span> <span class="s">$</span><span class="nv">\beta</span><span class="nb"> &gt; </span><span class="m">3</span><span class="s">$</span></div><div class='line' id='LC227'><span class="k">\\</span> <span class="k">\hline</span></div><div class='line' id='LC228'>glts, elts, glts, <span class="nb">&amp;</span> ilts <span class="nb">&amp;</span> 70<span class="k">\%</span> <span class="nb">&amp;</span> <span class="s">$</span><span class="nv">\beta</span><span class="nb"> &gt; </span><span class="m">5</span><span class="s">$</span></div><div class='line' id='LC229'><span class="k">\\</span> flts</div><div class='line' id='LC230'><span class="k">\\</span> <span class="k">\hline</span></div><div class='line' id='LC231'>ab, abb <span class="nb">&amp;</span> abbb <span class="nb">&amp;</span> 90<span class="k">\%</span> <span class="nb">&amp;</span> <span class="s">$</span><span class="nv">\alpha</span><span class="nb"> </span><span class="o">=</span><span class="nb"> </span><span class="m">0</span><span class="nb">, </span><span class="nv">\beta</span><span class="nb"> </span><span class="o">=</span><span class="nb"> </span><span class="m">4</span><span class="s">$</span></div><div class='line' id='LC232'><span class="k">\\</span> <span class="k">\hline</span></div><div class='line' id='LC233'>ao, aoo, aooo <span class="nb">&amp;</span> aoooo <span class="nb">&amp;</span> 80<span class="k">\%</span> <span class="nb">&amp;</span> <span class="s">$</span><span class="nv">\alpha</span><span class="nb"> </span><span class="o">=</span><span class="nb"> </span><span class="m">0</span><span class="nb">, </span><span class="nv">\beta</span><span class="o">=</span><span class="m">4</span><span class="s">$</span></div><div class='line' id='LC234'><span class="k">\\</span> <span class="k">\hline</span></div><div class='line' id='LC235'><span class="k">\end</span><span class="nb">{</span>tabular<span class="nb">}</span> </div><div class='line' id='LC236'><span class="k">\end</span><span class="nb">{</span>center<span class="nb">}</span> </div><div class='line' id='LC237'><span class="k">\end</span><span class="nb">{</span>table<span class="nb">}</span></div><div class='line' id='LC238'><br/></div><div class='line' id='LC239'><span class="k">\subsection</span><span class="nb">{</span>The Regex Game<span class="nb">}</span></div><div class='line' id='LC240'>Given the fitted model, we then considered the following task: </div><div class='line' id='LC241'><br/></div><div class='line' id='LC242'>Given a language <span class="s">$</span><span class="nv">\mathcal</span><span class="nb">{L}</span><span class="s">$</span>, a set of regular expressions <span class="s">$</span><span class="nv">\mathcal</span><span class="nb">{H}</span><span class="s">$</span> (where these are candidates for <span class="s">$</span><span class="nv">\mathcal</span><span class="nb">{L}</span><span class="s">$</span>), a set of strings drawn <span class="s">$</span><span class="nb">S</span><span class="s">$</span> from <span class="s">$</span><span class="nv">\mathcal</span><span class="nb">{L}</span><span class="s">$</span>, and a string <span class="s">$</span><span class="nb">s</span><span class="s">$</span>, we define</div><div class='line' id='LC243'><br/></div><div class='line' id='LC244'><span class="k">\begin</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC245'>	p(s <span class="k">\in</span> <span class="k">\mathcal</span><span class="nb">{</span>L<span class="nb">}</span> | S <span class="k">\subseteq</span> <span class="nb">{</span><span class="k">\mathcal</span><span class="nb">{</span>L<span class="nb">}}</span>) = <span class="k">\sum</span><span class="nb">_{</span>h <span class="k">\in</span> <span class="k">\mathcal</span><span class="nb">{</span>H<span class="nb">}}</span> 1 <span class="k">\{</span>s <span class="k">\in</span> h<span class="k">\}</span> p(h|S)</div><div class='line' id='LC246'><span class="k">\end</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC247'><br/></div><div class='line' id='LC248'>For human participants, because the hypothesis is unobserved, we take <span class="s">$</span><span class="nb">p</span><span class="o">(</span><span class="nb">s </span><span class="nv">\in</span><span class="nb"> </span><span class="nv">\mathcal</span><span class="nb">{L} | S </span><span class="nv">\subseteq</span><span class="nb"> {</span><span class="nv">\mathcal</span><span class="nb">{L}}</span><span class="o">)</span><span class="nb"> </span><span class="s">$</span> to be the average acceptance of a test string given a set <span class="s">$</span><span class="nb">S</span><span class="s">$</span>. </div><div class='line' id='LC249'><br/></div><div class='line' id='LC250'>We define <span class="s">$</span><span class="nv">\mathcal</span><span class="nb">{H}</span><span class="s">$</span> to be the set of all hypotheses generated in our model. Note, however, that this is the set of regular languages which were identified in beam search, so it does not represent the set of all possible hypotheses.</div><div class='line' id='LC251'><br/></div><div class='line' id='LC252'><span class="k">\section</span><span class="nb">{</span>Results and Discussion<span class="nb">}</span></div><div class='line' id='LC253'>Figure~<span class="k">\ref</span><span class="nb">{</span>results-chart<span class="nb">}</span> shows the comparison between our model and the human subject data.</div><div class='line' id='LC254'><br/></div><div class='line' id='LC255'><span class="k">\begin</span><span class="nb">{</span>figure<span class="nb">}</span>[ht]</div><div class='line' id='LC256'><span class="k">\begin</span><span class="nb">{</span>center<span class="nb">}</span></div><div class='line' id='LC257'><span class="k">\begin</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC258'><span class="k">\includegraphics</span><span class="na">[scale=0.6]</span><span class="nb">{</span>results.png<span class="nb">}</span></div><div class='line' id='LC259'><span class="k">\end</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC260'><span class="k">\end</span><span class="nb">{</span>center<span class="nb">}</span></div><div class='line' id='LC261'><span class="k">\caption</span><span class="nb">{</span>Chart showing  <span class="s">$</span><span class="nb">p</span><span class="o">(</span><span class="nb">s </span><span class="nv">\in</span><span class="nb"> </span><span class="nv">\mathcal</span><span class="nb">{L} | S </span><span class="nv">\subseteq</span><span class="nb"> {L}</span><span class="o">)</span><span class="nb"> </span><span class="s">$</span>. Human subject data is in red, model data is in blue. <span class="s">$</span><span class="nb">R</span><span class="s">$</span> = .75<span class="nb">}</span> </div><div class='line' id='LC262'><span class="k">\label</span><span class="nb">{</span>results-chart<span class="nb">}</span></div><div class='line' id='LC263'><span class="k">\end</span><span class="nb">{</span>figure<span class="nb">}</span> </div><div class='line' id='LC264'><br/></div><div class='line' id='LC265'>We note that subjects completed the 10 or 11 question survey on average in around three minutes. It remains unclear if more standardized results would have emerged in certain cases if subjects had taken more time with the task. In addition, more fine grained results may have been obtained if we had instead allowed users to bet certain amounts on whether or not a test string was in <span class="s">$</span><span class="nb">S</span><span class="s">$</span>. </div><div class='line' id='LC266'><br/></div><div class='line' id='LC267'>We note that in cases 1 (S = <span class="k">\{\verb</span>!yyyt!, <span class="k">\verb</span>!yt!, <span class="k">\verb</span>!yyyt!, <span class="k">\verb</span>!yyyt!, <span class="k">\verb</span>!yt!<span class="k">\}</span>, s= <span class="k">\verb</span>!yyyyyt!), 4 (S = <span class="k">\{\verb</span>!jojojojo!, <span class="k">\verb</span>!jojojojo!, <span class="k">\verb</span>!jojojojo!<span class="k">\}</span>, s = <span class="k">\verb</span>!jojo!), and 6 (S = <span class="k">\{\verb</span>!twtwtw!, <span class="k">\verb</span>!twtwtw!, <span class="k">\verb</span>!twtwtw!<span class="k">\}</span>, s =  <span class="k">\verb</span>!tw!), it is clear that the model weighs a smaller number of transitions greater than human subjects do. However, because this same parameter in part controls for wildcards, lowering it reduces performance on many other cases. This suggests that our model&#39;s parameter <span class="s">$</span><span class="nv">\beta</span><span class="s">$</span> may in fact be better represented by more than one variable in human decision making. </div><div class='line' id='LC268'><br/></div><div class='line' id='LC269'>We also observe that in such cases, there was a relatively small set of probable hypotheses, and that any favoring of smaller hypotheses was magnified by this fact. </div><div class='line' id='LC270'><span class="k">\section</span><span class="nb">{</span>Conclusion and Future Work<span class="nb">}</span></div><div class='line' id='LC271'>future work idea:</div><div class='line' id='LC272'><br/></div><div class='line' id='LC273'>-different wildcards (numbers, etc). this was implemented, but we didn&#39;t implement any experiments</div><div class='line' id='LC274'><br/></div><div class='line' id='LC275'>Despite the number of ways in which our model might be improved mentioned above,  it remains highly implausible that the brain generates its set of candidate hypotheses by ever considering even <span class="s">$</span><span class="m">2</span><span class="nb">^n</span><span class="s">$</span> hypotheses in the first level of beam search we perform, much less more than that. Doing so would be highly inefficient: the number of implausible hypotheses generated would in most cases be far greater than the number of `good&#39; hypotheses. Far more plausible a priori is the hypothesis that the brain uses of number of heuristics to generate a certain set of candidate hypothesis, and then uses some process analogous to Bayesian inference to select among those hypotheses. To shed light on how this hypothesis selection process may occur, we designed an experiment in which candidates chose among a number of possible rules to explain a set of strings S, but have yet to gather enough data to reach any significant conclusions. </div><div class='line' id='LC276'><br/></div><div class='line' id='LC277'>Further experiments could be conducted to determine what heuristics may be at work in the hypothesis generation process. For any set of candidate heuristics, it will remain to be determined whether or not such heuristics are at some point learned. Clearly some semantic considerations are at work in human pattern recognition on strings: cat and mat but not qat are meaningful english words; `ao&#39; and `eo&#39; and `io&#39; begin with vowels unlike `xo&#39;. (This experiment was designed to avoid such candidate hypotheses.)</div><div class='line' id='LC278'><br/></div><div class='line' id='LC279'>We note that we do not directly solve the problem of finding a most plausible way to align a set of strings <span class="s">$</span><span class="nb">S</span><span class="s">$</span>. Humans are excellent at quickly noticing the pattern in `aall&#39;, `ball&#39;, and `call&#39;. Although our model will infer that the last three letters can be represented with only three states in each case, in doing so it generates many candidate regular expressions which a human being would not consciously consider.  </div><div class='line' id='LC280'><br/></div><div class='line' id='LC281'>Future work may be directed towards investigating the hypothesis that humans first align candidate strings and then generate a number of hypotheses. For example, humans would be very quick to notice that ``shrew&quot;, ``reward&quot;, and ``threw&quot; all contain the `rew&#39; subsequence, and would most likely generate candidate regular expressions which preserved that substring. However, following <span class="k">\citeA</span><span class="nb">{</span>Fernau:2009:ALR:1512988.1513048<span class="nb">}</span> we note that the related task of the multiple sequence alignment problem is NP hard, and that a similar model which considers the most likely alignment of a set of strings given many possible alignments will consider a very large number of hypotheses when strings are long enough.  </div><div class='line' id='LC282'><br/></div><div class='line' id='LC283'><br/></div><div class='line' id='LC284'><br/></div><div class='line' id='LC285'><br/></div><div class='line' id='LC286'><span class="c">% \section{General Formatting Instructions}</span></div><div class='line' id='LC287'><br/></div><div class='line' id='LC288'><span class="c">% For standard spoken papers and standard posters, the entire contribution (including figures, references, everything) can be no longer than six pages. For abstract posters, the entire contribution can be no longer than one page. For symposia, the entire contribution can be no longer than two pages.</span></div><div class='line' id='LC289'><br/></div><div class='line' id='LC290'><span class="c">% The text of the paper should be formatted in two columns with an</span></div><div class='line' id='LC291'><span class="c">% overall width of 7 inches (17.8 cm) and length of 9.25 inches (23.5</span></div><div class='line' id='LC292'><span class="c">% cm), with 0.25 inches between the columns. Leave two line spaces</span></div><div class='line' id='LC293'><span class="c">% between the last author listed and the text of the paper. The left</span></div><div class='line' id='LC294'><span class="c">% margin should be 0.75 inches and the top margin should be 1 inch.</span></div><div class='line' id='LC295'><span class="c">% \textbf{The right and bottom margins will depend on whether you use</span></div><div class='line' id='LC296'><span class="c">%   U.S. letter or A4 paper, so you must be sure to measure the width of</span></div><div class='line' id='LC297'><span class="c">%   the printed text.} Use 10~point Times Roman with 12~point vertical</span></div><div class='line' id='LC298'><span class="c">% spacing, unless otherwise specified.</span></div><div class='line' id='LC299'><br/></div><div class='line' id='LC300'><span class="c">% The title should be in 14~point, bold, and centered. The title should</span></div><div class='line' id='LC301'><span class="c">% be formatted with initial caps (the first letter of content words</span></div><div class='line' id='LC302'><span class="c">% capitalized and the rest lower case). Each author&#39;s name should appear</span></div><div class='line' id='LC303'><span class="c">% on a separate line, 11~point bold, and centered, with the author&#39;s</span></div><div class='line' id='LC304'><span class="c">% email address in parentheses. Under each author&#39;s name list the</span></div><div class='line' id='LC305'><span class="c">% author&#39;s affiliation and postal address in ordinary 10~point type.</span></div><div class='line' id='LC306'><br/></div><div class='line' id='LC307'><span class="c">% Indent the first line of each paragraph by 1/8~inch (except for the</span></div><div class='line' id='LC308'><span class="c">% first paragraph of a new section). Do not add extra vertical space</span></div><div class='line' id='LC309'><span class="c">% between paragraphs.</span></div><div class='line' id='LC310'><br/></div><div class='line' id='LC311'><br/></div><div class='line' id='LC312'><span class="c">% \section{First-Level Headings}</span></div><div class='line' id='LC313'><br/></div><div class='line' id='LC314'><span class="c">% First level headings should be in 12~point, initial caps, bold and</span></div><div class='line' id='LC315'><span class="c">% centered. Leave one line space above the heading and 1/4~ line space</span></div><div class='line' id='LC316'><span class="c">% below the heading.</span></div><div class='line' id='LC317'><br/></div><div class='line' id='LC318'><br/></div><div class='line' id='LC319'><span class="c">% \subsection{Second-Level Headings}</span></div><div class='line' id='LC320'><br/></div><div class='line' id='LC321'><span class="c">% Second level headings should be 11~point, initial caps, bold, and</span></div><div class='line' id='LC322'><span class="c">% flush left. Leave one line space above the heading and 1/4~ line</span></div><div class='line' id='LC323'><span class="c">% space below the heading.</span></div><div class='line' id='LC324'><br/></div><div class='line' id='LC325'><br/></div><div class='line' id='LC326'><span class="c">% \subsubsection{Third-Level Headings}</span></div><div class='line' id='LC327'><br/></div><div class='line' id='LC328'><span class="c">% Third-level headings should be 10~point, initial caps, bold, and flush</span></div><div class='line' id='LC329'><span class="c">% left. Leave one line space above the heading, but no space after the</span></div><div class='line' id='LC330'><span class="c">% heading.</span></div><div class='line' id='LC331'><br/></div><div class='line' id='LC332'><br/></div><div class='line' id='LC333'><span class="c">% \section{Formalities, Footnotes, and Floats}</span></div><div class='line' id='LC334'><br/></div><div class='line' id='LC335'><span class="c">% Use standard APA citation format. Citations within the text should</span></div><div class='line' id='LC336'><span class="c">% include the author&#39;s last name and year. If the authors&#39; names are</span></div><div class='line' id='LC337'><span class="c">% included in the sentence, place only the year in parentheses, as in</span></div><div class='line' id='LC338'><span class="c">% \citeA{NewellSimon1972a}, but otherwise place the entire reference in</span></div><div class='line' id='LC339'><span class="c">% parentheses with the authors and year separated by a comma</span></div><div class='line' id='LC340'><span class="c">% \cite{NewellSimon1972a}. List multiple references alphabetically and</span></div><div class='line' id='LC341'><span class="c">% separate them by semicolons</span></div><div class='line' id='LC342'><span class="c">% \cite{ChalnickBillman1988a,NewellSimon1972a}. Use the</span></div><div class='line' id='LC343'><span class="c">% et~al. construction only after listing all the authors to a</span></div><div class='line' id='LC344'><span class="c">% publication in an earlier reference and for citations with four or</span></div><div class='line' id='LC345'><span class="c">% more authors.</span></div><div class='line' id='LC346'><br/></div><div class='line' id='LC347'><br/></div><div class='line' id='LC348'><span class="c">% \subsection{Footnotes}</span></div><div class='line' id='LC349'><br/></div><div class='line' id='LC350'><span class="c">% Indicate footnotes with a number\footnote{Sample of the first</span></div><div class='line' id='LC351'><span class="c">%   footnote.} in the text. Place the footnotes in 9~point type at the</span></div><div class='line' id='LC352'><span class="c">% bottom of the page on which they appear. Precede the footnote with a</span></div><div class='line' id='LC353'><span class="c">% horizontal rule.\footnote{Sample of the second footnote.}</span></div><div class='line' id='LC354'><br/></div><div class='line' id='LC355'><br/></div><div class='line' id='LC356'><span class="c">% \subsection{Tables}</span></div><div class='line' id='LC357'><br/></div><div class='line' id='LC358'><span class="c">% Number tables consecutively; place the table number and title (in</span></div><div class='line' id='LC359'><span class="c">% 10~point) above the table with one line space above the caption and</span></div><div class='line' id='LC360'><span class="c">% one line space below it, as in Table~\ref{sample-table}. You may float</span></div><div class='line' id='LC361'><span class="c">% tables to the top or bottom of a column, set wide tables across both</span></div><div class='line' id='LC362'><span class="c">% columns.</span></div><div class='line' id='LC363'><br/></div><div class='line' id='LC364'><span class="c">% \begin{table}[!ht]</span></div><div class='line' id='LC365'><span class="c">% \begin{center} </span></div><div class='line' id='LC366'><span class="c">% \caption{Sample table title.} </span></div><div class='line' id='LC367'><span class="c">% \label{sample-table} </span></div><div class='line' id='LC368'><span class="c">% \vskip 0.12in</span></div><div class='line' id='LC369'><span class="c">% \begin{tabular}{ll} </span></div><div class='line' id='LC370'><span class="c">% \hline</span></div><div class='line' id='LC371'><span class="c">% Error type    &amp;  Example \\</span></div><div class='line' id='LC372'><span class="c">% \hline</span></div><div class='line' id='LC373'><span class="c">% Take smaller        &amp;   63 - 44 = 21 \\</span></div><div class='line' id='LC374'><span class="c">% Always borrow~~~~   &amp;   96 - 42 = 34 \\</span></div><div class='line' id='LC375'><span class="c">% 0 - N = N           &amp;   70 - 47 = 37 \\</span></div><div class='line' id='LC376'><span class="c">% 0 - N = 0           &amp;   70 - 47 = 30 \\</span></div><div class='line' id='LC377'><span class="c">% \hline</span></div><div class='line' id='LC378'><span class="c">% \end{tabular} </span></div><div class='line' id='LC379'><span class="c">% \end{center} </span></div><div class='line' id='LC380'><span class="c">% \end{table}</span></div><div class='line' id='LC381'><br/></div><div class='line' id='LC382'><br/></div><div class='line' id='LC383'><span class="c">% \subsection{Figures}</span></div><div class='line' id='LC384'><br/></div><div class='line' id='LC385'><span class="c">% All artwork must be very dark for purposes of reproduction and should</span></div><div class='line' id='LC386'><span class="c">% not be hand drawn. Number figures sequentially, placing the figure</span></div><div class='line' id='LC387'><span class="c">% number and caption, in 10~point, after the figure with one line space</span></div><div class='line' id='LC388'><span class="c">% above the caption and one line space below it, as in</span></div><div class='line' id='LC389'><span class="c">% Figure~\ref{sample-figure}. If necessary, leave extra white space at</span></div><div class='line' id='LC390'><span class="c">% the bottom of the page to avoid splitting the figure and figure</span></div><div class='line' id='LC391'><span class="c">% caption. You may float figures to the top or bottom of a column, or</span></div><div class='line' id='LC392'><span class="c">% set wide figures across both columns.</span></div><div class='line' id='LC393'><br/></div><div class='line' id='LC394'><span class="c">% \begin{figure}[ht]</span></div><div class='line' id='LC395'><span class="c">% \begin{center}</span></div><div class='line' id='LC396'><span class="c">% \fbox{CoGNiTiVe ScIeNcE}</span></div><div class='line' id='LC397'><span class="c">% \end{center}</span></div><div class='line' id='LC398'><span class="c">% \caption{This is a figure.} </span></div><div class='line' id='LC399'><span class="c">% \label{sample-figure}</span></div><div class='line' id='LC400'><span class="c">% \end{figure}</span></div><div class='line' id='LC401'><br/></div><div class='line' id='LC402'><br/></div><div class='line' id='LC403'><span class="c">% \section{Acknowledgments}</span></div><div class='line' id='LC404'><br/></div><div class='line' id='LC405'><span class="c">% Place acknowledgments (including funding information) in a section at</span></div><div class='line' id='LC406'><span class="c">% the end of the paper.</span></div><div class='line' id='LC407'><br/></div><div class='line' id='LC408'><br/></div><div class='line' id='LC409'><span class="c">% \section{References Instructions}</span></div><div class='line' id='LC410'><br/></div><div class='line' id='LC411'><span class="c">% Follow the APA Publication Manual for citation format, both within the</span></div><div class='line' id='LC412'><span class="c">% text and in the reference list, with the following exceptions: (a) do</span></div><div class='line' id='LC413'><span class="c">% not cite the page numbers of any book, including chapters in edited</span></div><div class='line' id='LC414'><span class="c">% volumes; (b) use the same format for unpublished references as for</span></div><div class='line' id='LC415'><span class="c">% published ones. Alphabetize references by the surnames of the authors,</span></div><div class='line' id='LC416'><span class="c">% with single author entries preceding multiple author entries. Order</span></div><div class='line' id='LC417'><span class="c">% references by the same authors by the year of publication, with the</span></div><div class='line' id='LC418'><span class="c">% earliest first.</span></div><div class='line' id='LC419'><br/></div><div class='line' id='LC420'><span class="c">% Use a first level section heading for the reference list. Use a</span></div><div class='line' id='LC421'><span class="c">% hanging indent style, with the first line of the reference flush</span></div><div class='line' id='LC422'><span class="c">% against the left margin and subsequent lines indented by 1/8~inch.</span></div><div class='line' id='LC423'><span class="c">% Below are example references for a conference paper, book chapter,</span></div><div class='line' id='LC424'><span class="c">% journal article, technical report, dissertation, book, and edited</span></div><div class='line' id='LC425'><span class="c">% volume, respectively.</span></div><div class='line' id='LC426'><br/></div><div class='line' id='LC427'><br/></div><div class='line' id='LC428'><span class="k">\bibliographystyle</span><span class="nb">{</span>apacite<span class="nb">}</span></div><div class='line' id='LC429'><span class="k">\setlength</span><span class="nb">{</span><span class="k">\bibleftmargin</span><span class="nb">}{</span>.125in<span class="nb">}</span></div><div class='line' id='LC430'><span class="k">\setlength</span><span class="nb">{</span><span class="k">\bibindent</span><span class="nb">}{</span>-<span class="k">\bibleftmargin</span><span class="nb">}</span></div><div class='line' id='LC431'>&nbsp;<span class="k">\bibliography</span><span class="nb">{</span>regexbib<span class="nb">}</span></div><div class='line' id='LC432'><br/></div><div class='line' id='LC433'><span class="k">\section</span><span class="nb">{</span>Appendix: Technical Notes<span class="nb">}</span></div><div class='line' id='LC434'><span class="k">\subsection</span><span class="nb">{</span>Regex and DFA Notation<span class="nb">}</span></div><div class='line' id='LC435'>We used a limited alphabet wherein normal characters are all lower case. We also used the following rules to define regular expressions.</div><div class='line' id='LC436'><span class="k">\begin</span><span class="nb">{</span>enumerate<span class="nb">}</span></div><div class='line' id='LC437'>	<span class="k">\item</span> Acceptance of specific strings is represented as concatenated characters. E.g. <span class="k">\verb</span>!abc! accepts only <span class="k">\verb</span>!abc!.</div><div class='line' id='LC438'>	<span class="k">\item</span> Acceptance of disjunctive strings is represnted by <span class="k">\verb</span>![a|b]!. E.g. <span class="k">\verb</span>![abc|c]!accepts only <span class="k">\verb</span>!abc! and <span class="k">\verb</span>!c!.</div><div class='line' id='LC439'>	<span class="k">\item</span> Acceptance of any character is represented by the wildcard <span class="k">\verb</span>!A!. E.g. <span class="k">\verb</span>!aA! accepts <span class="k">\verb</span>!aa!, <span class="k">\verb</span>!ab!, <span class="k">\verb</span>!ac!, etc.</div><div class='line' id='LC440'>	<span class="k">\item</span> Acceptance of the empty string is represented by <span class="k">\verb</span>!E!. <span class="k">\verb</span>!Ea! is the same as <span class="k">\verb</span>!a!.</div><div class='line' id='LC441'>	<span class="k">\item</span> Acceptance of any sequence repeated zero or more times is represented by the Kleene star, <span class="k">\verb</span>!a*!. E.g. <span class="k">\verb</span>!a*! accepts <span class="k">\verb</span>!E!, <span class="k">\verb</span>!a!, <span class="k">\verb</span>!aa!, etc, </div><div class='line' id='LC442'><span class="k">\end</span><span class="nb">{</span>enumerate<span class="nb">}</span></div><div class='line' id='LC443'><br/></div><div class='line' id='LC444'><span class="k">\subsection</span><span class="nb">{</span>Sizing the Hypothesis Space<span class="nb">}</span></div><div class='line' id='LC445'>Ignoring wildcards and ignoring the mergings that require subset construction, each hypothesis can be defined as a set of mutually exclusive subsets of the states contained in the base hypothesis. For instance, given states <span class="s">$</span><span class="nb">s_</span><span class="m">1</span><span class="nb">, s_</span><span class="m">2</span><span class="nb">, s_</span><span class="m">3</span><span class="s">$</span>, the possible hypotheses are:</div><div class='line' id='LC446'>	<span class="k">\begin</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC447'>		<span class="nb">&amp;</span>s<span class="nb">_</span>1, s<span class="nb">_</span>2, s<span class="nb">_</span>3</div><div class='line' id='LC448'>		<span class="k">\\</span><span class="nb">&amp;</span><span class="k">\text</span><span class="nb">{</span>merge<span class="nb">}</span>(s<span class="nb">_</span>1, s<span class="nb">_</span>2), s<span class="nb">_</span>3</div><div class='line' id='LC449'>		<span class="k">\\</span><span class="nb">&amp;</span><span class="k">\text</span><span class="nb">{</span>merge<span class="nb">}</span>(s<span class="nb">_</span>1, s<span class="nb">_</span>3), s<span class="nb">_</span>2</div><div class='line' id='LC450'>		<span class="k">\\</span><span class="nb">&amp;</span><span class="k">\text</span><span class="nb">{</span>merge<span class="nb">}</span>(s<span class="nb">_</span>2, s<span class="nb">_</span>3), s<span class="nb">_</span>1</div><div class='line' id='LC451'>		<span class="k">\\</span><span class="nb">&amp;</span><span class="k">\text</span><span class="nb">{</span>merge<span class="nb">}</span>(s<span class="nb">_</span>1, s<span class="nb">_</span>2, s<span class="nb">_</span>3)</div><div class='line' id='LC452'>	<span class="k">\end</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC453'>To generalize this to a case of <span class="s">$</span><span class="nb">n_b</span><span class="s">$</span> states in the base hypothesis, one can consider individually the number of regexes with <span class="s">$</span><span class="nb">n_f</span><span class="s">$</span> states in the finished regex. For the cases where <span class="s">$</span><span class="nb">n_b</span><span class="o">=</span><span class="nb">n_f</span><span class="s">$</span> or <span class="s">$</span><span class="nb">n_f</span><span class="o">=</span><span class="m">1</span><span class="s">$</span>, it is easy to verify that there exist only one regex. When creating hypotheses where <span class="s">$</span><span class="m">1</span><span class="nb"> &lt; n_f &lt; n_b</span><span class="s">$</span>, however, one can simply look to the case of <span class="s">$</span><span class="nb">n_b </span><span class="o">-</span><span class="nb"> </span><span class="m">1</span><span class="s">$</span>. In particular, for every <span class="s">$</span><span class="nb">n_f</span><span class="s">$</span>, one can duplicate all of the hypotheses from the <span class="s">$</span><span class="nb">n_b </span><span class="o">-</span><span class="nb"> </span><span class="m">1</span><span class="s">$</span> case with an additional unmerged state <span class="s">$</span><span class="nb">s_{n_b}</span><span class="s">$</span>. Further, one can add a new hypothesis for each subset of the hypotheses in the <span class="s">$</span><span class="nb">n_s</span><span class="o">-</span><span class="m">1</span><span class="nb">, n_f</span><span class="s">$</span> case. So, the expression for the total number of regexes given the number of states in the base hypothesis is given by <span class="s">$</span><span class="nb">g</span><span class="o">(</span><span class="nb">n_b</span><span class="o">)</span><span class="s">$</span>.</div><div class='line' id='LC454'>	<span class="k">\begin</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC455'>	f(a, b) <span class="nb">&amp;</span>= <span class="k">\left\{</span></div><div class='line' id='LC456'>			<span class="k">\begin</span><span class="nb">{</span>array<span class="nb">}{</span>cl<span class="nb">}</span></div><div class='line' id='LC457'>&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;		1 <span class="nb">&amp;</span> a=b <span class="k">\vee</span> b=1 <span class="k">\\</span></div><div class='line' id='LC458'>&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;		f(a-1, b-1) + b f(a-1, b) <span class="nb">&amp;</span> <span class="k">\mbox</span><span class="nb">{</span>otherwise<span class="nb">}</span></div><div class='line' id='LC459'>&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;<span class="k">\end</span><span class="nb">{</span>array<span class="nb">}</span><span class="k">\right</span>.</div><div class='line' id='LC460'>		<span class="k">\\</span> g(n<span class="nb">_</span>b) <span class="nb">&amp;</span>=<span class="k">\sum</span><span class="nb">_{</span>n<span class="nb">_</span>f=1<span class="nb">}^{</span>n<span class="nb">_</span>b<span class="nb">}</span> f(n<span class="nb">_</span>b, n<span class="nb">_</span>f)</div><div class='line' id='LC461'>	<span class="k">\end</span><span class="nb">{</span>align*<span class="nb">}</span></div><div class='line' id='LC462'>One of our experiment examples contained 25 nodes in the base hypothesis, which means that the hypothesis space is <span class="s">$</span><span class="nb">g</span><span class="o">(</span><span class="m">25</span><span class="o">)</span><span class="nv">\approx</span><span class="nb"> </span><span class="m">5</span><span class="nb">.</span><span class="m">0</span><span class="nb"> </span><span class="nv">\times</span><span class="nb"> </span><span class="m">10</span><span class="nb">^{</span><span class="m">17</span><span class="nb">}</span><span class="s">$</span> regular expressions. Needless to say, it is intractable to exhaustively search the hypothesis space.</div><div class='line' id='LC463'><br/></div><div class='line' id='LC464'><span class="k">\end</span><span class="nb">{</span>document<span class="nb">}</span></div></pre></div>
          </td>
        </tr>
      </table>
  </div>

          </div>
        </div>
      </div>

      <a href="#jump-to-line" rel="facebox" data-hotkey="l" class="js-jump-to-line" style="display:none">Jump to Line</a>
      <div id="jump-to-line" style="display:none">
        <h2>Jump to Line</h2>
        <form accept-charset="UTF-8" class="js-jump-to-line-form">
          <input class="textfield js-jump-to-line-field" type="text">
          <div class="full-button">
            <button type="submit" class="classy">
              Go
            </button>
          </div>
        </form>
      </div>

    </div>
</div>

<div id="js-frame-loading-template" class="frame frame-loading large-loading-area" style="display:none;">
  <img class="js-frame-loading-spinner" src="https://a248.e.akamai.net/assets.github.com/images/spinners/octocat-spinner-128.gif?1347543527" height="64" width="64">
</div>

        </div>
      </div>
      <div class="context-overlay"></div>
    </div>

      <div id="footer-push"></div><!-- hack for sticky footer -->
    </div><!-- end of wrapper - hack for sticky footer -->

      <!-- footer -->
      <div id="footer">
  <div class="container clearfix">

      <dl class="footer_nav">
        <dt>GitHub</dt>
        <dd><a href="https://github.com/about">About us</a></dd>
        <dd><a href="https://github.com/blog">Blog</a></dd>
        <dd><a href="https://github.com/contact">Contact &amp; support</a></dd>
        <dd><a href="http://enterprise.github.com/">GitHub Enterprise</a></dd>
        <dd><a href="http://status.github.com/">Site status</a></dd>
      </dl>

      <dl class="footer_nav">
        <dt>Applications</dt>
        <dd><a href="http://mac.github.com/">GitHub for Mac</a></dd>
        <dd><a href="http://windows.github.com/">GitHub for Windows</a></dd>
        <dd><a href="http://eclipse.github.com/">GitHub for Eclipse</a></dd>
        <dd><a href="http://mobile.github.com/">GitHub mobile apps</a></dd>
      </dl>

      <dl class="footer_nav">
        <dt>Services</dt>
        <dd><a href="http://get.gaug.es/">Gauges: Web analytics</a></dd>
        <dd><a href="http://speakerdeck.com">Speaker Deck: Presentations</a></dd>
        <dd><a href="https://gist.github.com">Gist: Code snippets</a></dd>
        <dd><a href="http://jobs.github.com/">Job board</a></dd>
      </dl>

      <dl class="footer_nav">
        <dt>Documentation</dt>
        <dd><a href="http://help.github.com/">GitHub Help</a></dd>
        <dd><a href="http://developer.github.com/">Developer API</a></dd>
        <dd><a href="http://github.github.com/github-flavored-markdown/">GitHub Flavored Markdown</a></dd>
        <dd><a href="http://pages.github.com/">GitHub Pages</a></dd>
      </dl>

      <dl class="footer_nav">
        <dt>More</dt>
        <dd><a href="http://training.github.com/">Training</a></dd>
        <dd><a href="https://github.com/edu">Students &amp; teachers</a></dd>
        <dd><a href="http://shop.github.com">The Shop</a></dd>
        <dd><a href="http://octodex.github.com/">The Octodex</a></dd>
      </dl>

      <hr class="footer-divider">

    <p class="right">&copy; 2012 <span title="0.16995s from fe2.rs.github.com">GitHub</span> Inc. All rights reserved.</p>
    <a class="left" href="https://github.com/">
      <span class="mega-icon mega-icon-invertocat"></span>
    </a>
    <ul id="legal">
        <li><a href="https://github.com/site/terms">Terms of Service</a></li>
        <li><a href="https://github.com/site/privacy">Privacy</a></li>
        <li><a href="https://github.com/security">Security</a></li>
    </ul>

  </div><!-- /.container -->

</div><!-- /.#footer -->

<div id="keyboard_shortcuts_pane" class="instapaper_ignore readability-extra" style="display:none">
  <h2>Keyboard Shortcuts <small><a href="#" class="js-see-all-keyboard-shortcuts">(see all)</a></small></h2>

  <div class="columns threecols">
    <div class="column first">
      <h3>Site wide shortcuts</h3>
      <dl class="keyboard-mappings">
        <dt>s</dt>
        <dd>Focus command bar</dd>
      </dl>
      <dl class="keyboard-mappings">
        <dt>?</dt>
        <dd>Bring up this help dialog</dd>
      </dl>
    </div><!-- /.column.first -->

    <div class="column middle" style='display:none'>
      <h3>Commit list</h3>
      <dl class="keyboard-mappings">
        <dt>j</dt>
        <dd>Move selection down</dd>
      </dl>
      <dl class="keyboard-mappings">
        <dt>k</dt>
        <dd>Move selection up</dd>
      </dl>
      <dl class="keyboard-mappings">
        <dt>c <em>or</em> o <em>or</em> enter</dt>
        <dd>Open commit</dd>
      </dl>
      <dl class="keyboard-mappings">
        <dt>y</dt>
        <dd>Expand URL to its canonical form</dd>
      </dl>
    </div><!-- /.column.first -->

    <div class="column last js-hidden-pane" style='display:none'>
      <h3>Pull request list</h3>
      <dl class="keyboard-mappings">
        <dt>j</dt>
        <dd>Move selection down</dd>
      </dl>
      <dl class="keyboard-mappings">
        <dt>k</dt>
        <dd>Move selection up</dd>
      </dl>
      <dl class="keyboard-mappings">
        <dt>o <em>or</em> enter</dt>
        <dd>Open issue</dd>
      </dl>
      <dl class="keyboard-mappings">
        <dt><span class="platform-mac">⌘</span><span class="platform-other">ctrl</span> <em>+</em> enter</dt>
        <dd>Submit comment</dd>
      </dl>
      <dl class="keyboard-mappings">
        <dt><span class="platform-mac">⌘</span><span class="platform-other">ctrl</span> <em>+</em> shift p</dt>
        <dd>Preview comment</dd>
      </dl>
    </div><!-- /.columns.last -->

  </div><!-- /.columns.equacols -->

  <div class="js-hidden-pane" style='display:none'>
    <div class="rule"></div>

    <h3>Issues</h3>

    <div class="columns threecols">
      <div class="column first">
        <dl class="keyboard-mappings">
          <dt>j</dt>
          <dd>Move selection down</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>k</dt>
          <dd>Move selection up</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>x</dt>
          <dd>Toggle selection</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>o <em>or</em> enter</dt>
          <dd>Open issue</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt><span class="platform-mac">⌘</span><span class="platform-other">ctrl</span> <em>+</em> enter</dt>
          <dd>Submit comment</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt><span class="platform-mac">⌘</span><span class="platform-other">ctrl</span> <em>+</em> shift p</dt>
          <dd>Preview comment</dd>
        </dl>
      </div><!-- /.column.first -->
      <div class="column last">
        <dl class="keyboard-mappings">
          <dt>c</dt>
          <dd>Create issue</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>l</dt>
          <dd>Create label</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>i</dt>
          <dd>Back to inbox</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>u</dt>
          <dd>Back to issues</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>/</dt>
          <dd>Focus issues search</dd>
        </dl>
      </div>
    </div>
  </div>

  <div class="js-hidden-pane" style='display:none'>
    <div class="rule"></div>

    <h3>Issues Dashboard</h3>

    <div class="columns threecols">
      <div class="column first">
        <dl class="keyboard-mappings">
          <dt>j</dt>
          <dd>Move selection down</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>k</dt>
          <dd>Move selection up</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>o <em>or</em> enter</dt>
          <dd>Open issue</dd>
        </dl>
      </div><!-- /.column.first -->
    </div>
  </div>

  <div class="js-hidden-pane" style='display:none'>
    <div class="rule"></div>

    <h3>Network Graph</h3>
    <div class="columns equacols">
      <div class="column first">
        <dl class="keyboard-mappings">
          <dt><span class="badmono">←</span> <em>or</em> h</dt>
          <dd>Scroll left</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt><span class="badmono">→</span> <em>or</em> l</dt>
          <dd>Scroll right</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt><span class="badmono">↑</span> <em>or</em> k</dt>
          <dd>Scroll up</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt><span class="badmono">↓</span> <em>or</em> j</dt>
          <dd>Scroll down</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>t</dt>
          <dd>Toggle visibility of head labels</dd>
        </dl>
      </div><!-- /.column.first -->
      <div class="column last">
        <dl class="keyboard-mappings">
          <dt>shift <span class="badmono">←</span> <em>or</em> shift h</dt>
          <dd>Scroll all the way left</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>shift <span class="badmono">→</span> <em>or</em> shift l</dt>
          <dd>Scroll all the way right</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>shift <span class="badmono">↑</span> <em>or</em> shift k</dt>
          <dd>Scroll all the way up</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>shift <span class="badmono">↓</span> <em>or</em> shift j</dt>
          <dd>Scroll all the way down</dd>
        </dl>
      </div><!-- /.column.last -->
    </div>
  </div>

  <div class="js-hidden-pane" >
    <div class="rule"></div>
    <div class="columns threecols">
      <div class="column first js-hidden-pane" >
        <h3>Source Code Browsing</h3>
        <dl class="keyboard-mappings">
          <dt>t</dt>
          <dd>Activates the file finder</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>l</dt>
          <dd>Jump to line</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>w</dt>
          <dd>Switch branch/tag</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>y</dt>
          <dd>Expand URL to its canonical form</dd>
        </dl>
      </div>
    </div>
  </div>

  <div class="js-hidden-pane" style='display:none'>
    <div class="rule"></div>
    <div class="columns threecols">
      <div class="column first">
        <h3>Browsing Commits</h3>
        <dl class="keyboard-mappings">
          <dt><span class="platform-mac">⌘</span><span class="platform-other">ctrl</span> <em>+</em> enter</dt>
          <dd>Submit comment</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>escape</dt>
          <dd>Close form</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>p</dt>
          <dd>Parent commit</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>o</dt>
          <dd>Other parent commit</dd>
        </dl>
      </div>
    </div>
  </div>

  <div class="js-hidden-pane" style='display:none'>
    <div class="rule"></div>
    <h3>Notifications</h3>

    <div class="columns threecols">
      <div class="column first">
        <dl class="keyboard-mappings">
          <dt>j</dt>
          <dd>Move selection down</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>k</dt>
          <dd>Move selection up</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>o <em>or</em> enter</dt>
          <dd>Open notification</dd>
        </dl>
      </div><!-- /.column.first -->

      <div class="column second">
        <dl class="keyboard-mappings">
          <dt>e <em>or</em> shift i <em>or</em> y</dt>
          <dd>Mark as read</dd>
        </dl>
        <dl class="keyboard-mappings">
          <dt>shift m</dt>
          <dd>Mute thread</dd>
        </dl>
      </div><!-- /.column.first -->
    </div>
  </div>

</div>

    <div id="markdown-help" class="instapaper_ignore readability-extra">
  <h2>Markdown Cheat Sheet</h2>

  <div class="cheatsheet-content">

  <div class="mod">
    <div class="col">
      <h3>Format Text</h3>
      <p>Headers</p>
      <pre>
# This is an &lt;h1&gt; tag
## This is an &lt;h2&gt; tag
###### This is an &lt;h6&gt; tag</pre>
     <p>Text styles</p>
     <pre>
*This text will be italic*
_This will also be italic_
**This text will be bold**
__This will also be bold__

*You **can** combine them*
</pre>
    </div>
    <div class="col">
      <h3>Lists</h3>
      <p>Unordered</p>
      <pre>
* Item 1
* Item 2
  * Item 2a
  * Item 2b</pre>
     <p>Ordered</p>
     <pre>
1. Item 1
2. Item 2
3. Item 3
   * Item 3a
   * Item 3b</pre>
    </div>
    <div class="col">
      <h3>Miscellaneous</h3>
      <p>Images</p>
      <pre>
![GitHub Logo](/images/logo.png)
Format: ![Alt Text](url)
</pre>
     <p>Links</p>
     <pre>
http://github.com - automatic!
[GitHub](http://github.com)</pre>
<p>Blockquotes</p>
     <pre>
As Kanye West said:

> We're living the future so
> the present is our past.
</pre>
    </div>
  </div>
  <div class="rule"></div>

  <h3>Code Examples in Markdown</h3>
  <div class="col">
      <p>Syntax highlighting with <a href="http://github.github.com/github-flavored-markdown/" title="GitHub Flavored Markdown" target="_blank">GFM</a></p>
      <pre>
```javascript
function fancyAlert(arg) {
  if(arg) {
    $.facebox({div:'#foo'})
  }
}
```</pre>
    </div>
    <div class="col">
      <p>Or, indent your code 4 spaces</p>
      <pre>
Here is a Python code example
without syntax highlighting:

    def foo:
      if not bar:
        return true</pre>
    </div>
    <div class="col">
      <p>Inline code for comments</p>
      <pre>
I think you should use an
`&lt;addr&gt;` element here instead.</pre>
    </div>
  </div>

  </div>
</div>

    <div id="ajax-error-message" class="flash flash-error">
      <span class="mini-icon mini-icon-exclamation"></span>
      Something went wrong with that request. Please try again.
      <a href="#" class="mini-icon mini-icon-remove-close ajax-error-dismiss"></a>
    </div>

    <span id='server_response_time' data-time='0.17114' data-host='fe2'></span>

  </body>
</html>